# Flow data

```{r}
#| label: dataDataByRiverFlowOptions
#| include: false
knitr::opts_chunk$set(message = FALSE, warning = FALSE)
```

```{r}
#| label: librariesDataFlowByRiver
#| echo: false

# for gams
library(gratia)
library(arm)
library(relaimpo)
library(GGally)
# library(linkedModels)
library(arrayhelpers)
library(mgcv)

library(getWBData)
library(lubridate)
library(kableExtra)
library(GGally)
library(lme4)
library(targets)
library(relaimpo)
library(caret)
library(pscl)
library(gridExtra)

library(getPrepareWBData) # for theme_publication()
library(ggthemes)

library(viridis)
library(tidyverse)

`%notin%` <- Negate(`%in%`)

source('./R/generalFunctions.R') #to get global_labeller_WB
```

```{r}
#| label: labellerModelsFlowByRiver
#| echo: false
# labelsIntSeason <- c(
#   "1" = "Spring",
#   "2" = "Summer",
#   "3" = "Autumn",
#   "4" = "Winter"
# )
# 
# labelsIntRiver <- c(
#   "west brook" = "West Brook",
#   "wb jimmy" = "Open Large",
#   "wb mitchell" = "Open Small",
#   "wb obear" = "Isolated Small"
# )
# 
# global_labellerInt <- labeller(
#   season = labelsIntSeason,
#   river = labelsIntRiver
#   #.default = label_both
# )
```

Read in fish and environmental data
```{r}
#| label: getDataGam
#| cache: false

library(targets) #Seems to be needed with cache: false

cd1 <- tar_read(cd1_target)  # from modelGrowthInMass_target
dGAM <- tar_read(dGAM_target) # from modelGrowthInMass_target

kIn = 4
speciesIn <- "Brook trout"

envData <- tar_read(envDataWB_target)
```

Read in flow duration curve data
```{r}
#| label: getFDC
#| cache: false
#| 
library(targets) #Seems to be needed with cache: false

fdcStats <- tar_read(fdcStats_target)
fdcStatsY <- tar_read(fdcStatsY_target)
fdcStatsS <- tar_read(fdcStatsS)
fdcStatsYS <- tar_read(fdcStatsYS_target)

envDataWB_fdcThresh <- tar_read(envDataWB_fdcThresh_target)
propFDC_aboveBelow <- tar_read(propFDC_aboveBelow_target)

```


## Raw daily flow data

Raw data by day.

Four types of flow data:\
\
Primary flow data types  
- *flowByArea_flowExt* (green), flow from the flow extension model scaled by watershed area as a proportion of WB area.\
- *FlowByRiver* (blue), flows for each river hindcast from Jenn's model based on tributary-specific flow estimates.\
\
Additional flow data types  
- *Flow* (red), flow extension model for the West Brook from the Mill River.\
- *FlowByArea_ByRiver* (orange), flow from the West Brook from Jenn's models scaled to tributaries based on relative tributary area.  

### Frequency plots of flow data types

```{r}
#| label: flow_envFlowFreq
#| fig-cap: "Stream flow distributions, type (red = flow, blue = flowByRiver, green = flowByArea_flowExt, orange = FlowByArea_ByRiver)"

ggplot(envData |> filter(!is.na(season)), aes(flow)) +
  #geom_density(color = viridis(10)[1], adjust = 2.5) +
  geom_density(color = "red3", adjust = 2.5, size = 2) +
  geom_density(aes(flowByRiver), color = "blue3", adjust = 2.5) +
  geom_density(aes(flowByArea_flowExt), color = "green4", adjust = 2.5) +
  geom_density(aes(flowByArea_ByRiver), color = "orange2", adjust = 2.5) +
  labs(x = "Flow (m3/s)") +
  scale_x_log10() +
  facet_grid(river~season, scales = 'free', labeller = global_labellerRiverSeasonInt_WB)
```

### Time series

::: panel-tabset
#### West Brook
```{r}
#| label: flowRawWB
#| fig-cap: "West Brook stream flow distributions, type (red = flow, blue = flowByRiver, green = flowByArea_flowExt, orange = FlowByArea_ByRiver)"

ggplot(envData |> filter(river == "west brook", year > 1999), aes(yday, flow)) +
  #geom_point(color = "red3") +
  geom_point(aes(yday, flowByRiver), color = "blue3", size = 0.1) +
  geom_point(aes(yday, flowByArea_flowExt), color = "green4", size = 0.5) +
  geom_point(aes(yday, flowByArea_ByRiver), color = "orange2", size = 0.5) +
  scale_y_log10() +
  facet_wrap(~year)

```

#### Open Large
```{r}
#| label: flowRawOL
#| fig-cap: "Open Large stream flow distributions, type (red = flow, blue = flowByRiver, green = flowByArea_flowExt, orange = FlowByArea_ByRiver)"

ggplot(envData |> filter(river == "wb jimmy", year > 1999), aes(yday, flow)) +
  #geom_point(color = "red3") +
  geom_point(aes(yday, flowByRiver), color = "blue3", size = 0.1) +
  geom_point(aes(yday, flowByArea_flowExt), color = "green4", size = 0.5) +
  geom_point(aes(yday, flowByArea_ByRiver), color = "orange2", size = 0.5) +
  scale_y_log10() +
  facet_wrap(~year)

```

#### Open Small
```{r}
#| label: flowRawOS
#| fig-cap: "Open Small stream flow distributions, type (red = flow, blue = flowByRiver, green = flowByArea_flowExt, orange = FlowByArea_ByRiver)"

ggplot(envData |> filter(river == "wb mitchell", year > 1999), aes(date, flow)) +
  #geom_line(color = "red3") +
  geom_point(aes(yday, flowByRiver), color = "blue3", size = 0.1) +
  geom_point(aes(yday, flowByArea_flowExt), color = "green4", size = 0.5) +
  geom_point(aes(yday, flowByArea_ByRiver), color = "orange2", size = 0.5) +
  scale_y_log10() +
  facet_wrap(~year)

```

#### Isolated Small
```{r}
#| label: flowRawIS
#| fig-cap: "Isolated Small stream flow distributions, type (red = flow, blue = flowByRiver, green = flowByArea_flowExt, orange = FlowByArea_ByRiver)"

ggplot(envData |> filter(river == "wb obear", year > 1999), aes(date, flow)) +
  #geom_line(color = "red3") +
  geom_point(aes(yday, flowByRiver), color = "blue3", size = 0.1) +
  geom_point(aes(yday, flowByArea_flowExt), color = "green4", size = 0.5) +
  geom_point(aes(yday, flowByArea_ByRiver), color = "orange2", size = 0.1) +
  scale_y_log10() +
  facet_wrap(~year)

```
:::
  
  #### Summary stats
```{r}
#| label: timeSeriesFlwoStats

summaryStats <- envData |> 
  filter(!is.na(season)) |> 
  group_by(river, season) |> 
  summarize(meanFlowByRiver = mean(flowByRiver, na.rm = TRUE),
            sdFlowByRiver = sd(flowByRiver, na.rm = TRUE),
            meanFlowByArea = mean(flowByArea_flowExt),
            sdFlowByArea = sd(flowByArea_flowExt))

summaryStatsLong <- 
  summaryStats |> 
  pivot_longer(cols = c(meanFlowByArea, meanFlowByRiver), names_to = "flowType") |> 
  addGG_noSpecies()

ggplot(summaryStatsLong, aes(seasonGG, value, color = flowType)) +
  geom_point(size = 4) +
  geom_line(aes(season,value, color = flowType), size = 1.2) + # using seasonGG in ggplot aes throws an error because it is an ordered factor
  scale_x_discrete("Season") +
  scale_y_continuous("Mean stream flow (m3/s)") +
  theme_publication() +
  #guides(color = FALSE) +
  facet_wrap(~riverGG, scales = "free")
```



### Flow duration curves

#### Across all years and seasons

90% for low flow - look for inflection point

Peak flow analysis for spring

```{r}
#| label: flowDurationCurves

ggplot(fdcStats, aes(stat, perc, color = variable)) +
  geom_line() +
  geom_point(size = 1) +
  labs(x = "Flow percentile", y = "Flow") +
  scale_y_log10() +
  facet_wrap(~river, scales = "free", labeller = global_labellerRiverSeasonInt_WB)
```

#### Grouped by year

::: panel-tabset
#### West Brook
```{r}
#| label: fdcWB

ggplot(fdcStatsY |> filter(river == "west brook"), aes(stat, perc, color = year, group = year)) +
  geom_line() +
  geom_point(size = 0.5) +
  labs(x = "Flow percentile", y = "Flow") +
  scale_y_log10() +
  scale_color_viridis() +
  facet_wrap(~variable, ncol = 2, labeller = global_labellerRiverSeasonInt_WB)
```

#### Open Large
```{r}
#| label: fdcOL

ggplot(fdcStatsY |> filter(river == "wb jimmy"), aes(stat, perc, color = year, group = year)) +
  geom_line() +
  geom_point(size = 0.5) +
  labs(x = "Flow percentile", y = "Flow") +
  scale_y_log10() +
  scale_color_viridis() +
  facet_wrap(~variable, ncol = 2, labeller = global_labellerRiverSeasonInt_WB)
```

#### Open Small
```{r}
#| label: fdcOS

ggplot(fdcStatsY |> filter(river == "wb mitchell"), aes(stat, perc, color = year, group = year)) +
  geom_line() +
  geom_point(size = 0.5) +
  labs(x = "Flow percentile", y = "Flow") +
  scale_y_log10() +
  scale_color_viridis() +
  facet_wrap(~variable, ncol = 2, labeller = global_labellerRiverSeasonInt_WB)
```

#### Isolated Small
```{r}
#| label: fdcIS

ggplot(fdcStatsY |> filter(river == "wb obear"), aes(stat, perc, color = year, group = year)) +
  geom_line() +
  geom_point(size = 0.5) +
  labs(x = "Flow percentile", y = "Flow") +
  scale_y_log10() +
  scale_color_viridis() +
  facet_wrap(~variable, ncol = 2, labeller = global_labellerRiverSeasonInt_WB)
```
:::
  
  #### Grouped by season
  
  ::: panel-tabset
#### West Brook
```{r}
#| label: fdcSWB

ggplot(fdcStatsS |> filter(river == "west brook", !is.na(season)), aes(stat, perc, color = factor(season))) +
  geom_line() +
  geom_point(size = 0.5) +
  labs(x = "Flow percentile", y = "Flow") +
  scale_y_log10() +
  facet_wrap(~variable, ncol = 2, labeller = global_labellerRiverSeasonInt_WB)
```

#### Open Large
```{r}
#| label: fdcSOL

ggplot(fdcStatsS |> filter(river == "wb jimmy", !is.na(season)), aes(stat, perc, color = factor(season))) +
  geom_line() +
  geom_point(size = 0.5) +
  labs(x = "Flow percentile", y = "Flow") +
  scale_y_log10() +
  facet_wrap(~variable, ncol = 2, labeller = global_labellerRiverSeasonInt_WB)
```

#### Open Small
```{r}
#| label: fdcSOS

ggplot(fdcStatsS |> filter(river == "wb mitchell", !is.na(season)), aes(stat, perc, color = factor(season))) +
  geom_line() +
  geom_point(size = 0.5) +
  labs(x = "Flow percentile", y = "Flow") +
  scale_y_log10() +
  facet_wrap(~variable, ncol = 2, labeller = global_labellerRiverSeasonInt_WB)
```

#### Isolated Small
```{r}
#| label: fdcSIS

ggplot(fdcStatsS |> filter(river == "wb obear", !is.na(season)), aes(stat, perc, color = factor(season))) +
  geom_line() +
  geom_point(size = 0.5) +
  labs(x = "Flow percentile", y = "Flow") +
  scale_y_log10() +
  facet_wrap(~variable, ncol = 2, labeller = global_labellerRiverSeasonInt_WB)
```
:::
  
  
  #### Grouped by year and season
  
  ::: panel-tabset
#### West Brook
```{r}
#| label: flowDurationCurvesYearSeasonWB

ggplot(fdcStatsYS |> filter(river == "west brook"), aes(stat, perc, color = year, group = year)) +
  geom_line() +
  geom_point(size = 0.5) +
  labs(x = "Flow percentile", y = "Flow") +
  scale_y_log10() +
  scale_color_viridis() +
  facet_grid(season~variable, scales = "free", labeller = global_labellerRiverSeasonInt_WB)
```

#### Open Large
```{r}
#| label: flowDurationCurvesYearSeasonOL

ggplot(fdcStatsYS |> filter(river == "wb jimmy"), aes(stat, perc, color = year, group = year)) +
  geom_line() +
  geom_point(size = 0.5) +
  labs(x = "Flow percentile", y = "Flow") +
  scale_y_log10() +
  scale_color_viridis() +
  facet_grid(season~variable, scales = "free", labeller = global_labellerRiverSeasonInt_WB)
```

#### Open Small
```{r}
#| label: flowDurationCurvesYearSeasonOS

ggplot(fdcStatsYS |> filter(river == "wb mitchell"), aes(stat, perc, color = year, group = year)) +
  geom_line() +
  geom_point(size = 0.5) +
  labs(x = "Flow percentile", y = "Flow") +
  scale_y_log10() +
  scale_color_viridis() +
  facet_grid(season~variable, scales = "free", labeller = global_labellerRiverSeasonInt_WB)
```

#### Isolated small
```{r}
#| label: flowDurationCurvesYearSeasonIS

ggplot(fdcStatsYS |> filter(river == "wb obear"), aes(stat, perc, color = year, group = year)) +
  geom_line() +
  geom_point(size = 0.5) +
  labs(x = "Flow percentile", y = "Flow") +
  scale_y_log10() +
  scale_color_viridis() +
  facet_grid(season~variable, scales = "free", labeller = global_labellerRiverSeasonInt_WB)
```
:::
  
  #### Proportion of days above/below flow duration curve thresholds  
  Thresholds defined in `fdcThresholds_target` in modelFDC_targets.R. Based on season-specific flow duration curves.  

::: panel-tabset
#### By river, high flows
```{r}
#| label: flowDurationCurvesBelowByRiver

ggplot(propFDC_aboveBelow |> filter(year > 1999), aes(year, propBelowLoFlowThreshByRiver, color = riverGG)) + 
  geom_point() +
  geom_line() +
  scale_color_discrete() +
  labs(x = "Year", y = "Proportion of days with flows higher than the high flow threshold") +
  facet_wrap(~season, labeller = global_labellerRiverSeasonInt_WB)

```

#### By river, low flows
```{r}
#| label: flowDurationCurvesAboveByRiver

ggplot(propFDC_aboveBelow |> filter(year > 1999), aes(year, propAboveHiFlowThreshRiver, color = riverGG)) + 
  geom_point() +
  geom_line() +
  scale_color_discrete() +
  labs(x = "Year", y = "Proportion of days with flows lower than the low flow threshold") +
  facet_wrap(~season, labeller = global_labellerRiverSeasonInt_WB)

```

#### By area, high flows
```{r}
#| label: flowDurationCurvesBelowByArea

ggplot(propFDC_aboveBelow |> filter(year > 1999), aes(year, propBelowLoFlowThreshByArea_flowExt, color = riverGG)) + 
  geom_point() +
  geom_line() +
  scale_color_discrete() +
  labs(x = "Year", y = "Proportion of days with flows higher than the high flow threshold") +
  facet_wrap(~season, labeller = global_labellerRiverSeasonInt_WB)

```

#### By area, low flows
```{r}
#| label: flowDurationCurvesAboveByArea

ggplot(propFDC_aboveBelow |> filter(year > 1999), aes(year, propAboveHiFlowThreshByArea_flowExt, color = riverGG)) + 
  geom_point() +
  geom_line() +
  scale_color_discrete() +
  labs(x = "Year", y = "Proportion of days with flows lower than the low flow threshold") +
  facet_wrap(~season, labeller = global_labellerRiverSeasonInt_WB)

```
:::
  #### Proportion of days above/below flow duration curve thresholds for individual fish data      
  Thresholds defined in `fdcThresholds_target` in modelFDC_targets.R. Based on season-specific flow duration curves. 
Individual-specific proportion of days calculated in `addEnvironmental3()` in `getDAtaElectro.R` and in `addEnvrionmental()` in `getDataCMR_targets.R`.  
This series of figures confirms that values from `cd1` are similar to seasonal values from `propFDC_aboveBelow` and also shows the indvidual variation in proportion of days stats.  

::: panel-tabset
#### By river, high flows
```{r}
#| label: flowDurationCurvescd1BelowByRiver

ggplot(cd1 |> filter(year > 1999), aes(year, propBelowLoFlowThreshByRiver, color = riverGG)) +
  geom_point() +
  geom_line() +
  scale_color_discrete() +
  labs(x = "Year", y = "Proportion of days with flows higher than the high flow threshold") +
  facet_wrap(~season, labeller = global_labellerRiverSeasonInt_WB)
```

#### By river, low flows
```{r}
#| label: flowDurationCurvescd1AboveByRiver

ggplot(cd1 |> filter(year > 1999), aes(year, propAboveHiFlowThreshByRiver, color = riverGG)) + 
  geom_point() +
  geom_line() +
  scale_color_discrete() +
  labs(x = "Year", y = "Proportion of days with flows lower than the low flow threshold") +
  facet_wrap(~season, labeller = global_labellerRiverSeasonInt_WB)

```

#### By area, high flows
```{r}
#| label: flowDurationCurvescd1BelowByArea

ggplot(cd1 |> filter(year > 1999), aes(year, propBelowLoFlowThreshByArea_flowExt, color = riverGG)) + 
  geom_point() +
  geom_line() +
  scale_color_discrete() +
  labs(x = "Year", y = "Proportion of days with flows higher than the high flow threshold") +
  facet_wrap(~season, labeller = global_labellerRiverSeasonInt_WB)

```

#### By area, low flows
```{r}
#| label: flowDurationCurvescd1AboveByArea

ggplot(cd1 |> filter(year > 1999), aes(year, propAboveHiFlowThreshByArea_flowExt, color = riverGG)) + 
  geom_point() +
  geom_line() +
  scale_color_discrete() +
  labs(x = "Year", y = "Proportion of days with flows lower than the low flow threshold") +
  facet_wrap(~season, labeller = global_labellerRiverSeasonInt_WB)

```
:::
  
  ### Raw *mean* flow data
  
  Mean flow data that go into models as independent data. Calculated for each fish based on dates of capture for each interval.  
#### Mean flow data types plotted against each other

```{r}
#| label: flow_flowByRiverRaw1
#| fig-cap: "Stream flow (m3/s) for the West Brook against stream flow by river"
#| 
ggplot(cd1, aes(meanFlow, meanFlowByRiver)) +
  geom_point() +
  geom_abline(slope = 1, intercept = 0) +
  facet_grid(river~season, scales = 'free', labeller = global_labellerRiverSeasonInt_WB)

```

```{r}
#| label: flow_flowByRiverRaw2
#| fig-cap: "Stream flow (m3/s) against stream flow by area"

ggplot(cd1, aes(meanFlow, meanFlowByArea_flowExt)) +
  geom_point() +
  geom_abline(slope = 1, intercept = 0) +
  facet_grid(river~season, scales = 'free', labeller = global_labellerRiverSeasonInt_WB)
```

```{r}
#| label: flow_flowByRiverRaw3
#| fig-cap: "Stream flow by river (m3/s) against stream flow by area"

ggplot(cd1, aes(meanFlowByRiver, meanFlowByArea_flowExt)) +
  geom_point() +
  geom_abline(slope = 1, intercept = 0) +
  facet_grid(river~season, scales = 'free', labeller = global_labellerRiverSeasonInt_WB)

```

#### Frequency plots of mean flow data types

```{r}
#| label: flow_flowByRiverRawFreqs3
#| fig-cap: "Stream flow distributions, type (red = flow, blue = flowByRiver, green = flowByArea_flowExt)"

ggplot(cd1, aes(meanFlow)) +
  #geom_density(color = viridis(10)[1], adjust = 2.5) +
  geom_density(color = "red3", adjust = 2.5) +
  geom_density(aes(meanFlowByRiver), color = "blue3", adjust = 2.5) +
  geom_density(aes(meanFlowByArea_flowExt), color = "green4", adjust = 2.5) +
  labs(x = "Flow m3/s") +
  facet_grid(river~season, scales = 'free', labeller = global_labellerRiverSeasonInt_WB)
```

#### Time series plots of flow data types

```{r}
#| label: flow_flowByRiverRawTimeSeries
#| fig-cap: "Stream flow time series, flow type (blue = flowByRiver, green = flowByArea_flowExt)"

ggplot(cd1) +
  #geom_point(aes(date, meanFlowByRiver), color = "blue3") +
  geom_line(aes(date, meanFlowByRiver), color = "blue3") +
  geom_line(aes(date, meanFlowByArea_flowExt), color = "green4", adjust = 2.5) +
  labs(x = "Date", y = "Stream flow") +
  facet_grid(river~season, scales = 'free', labeller = global_labellerRiverSeasonInt_WB)
```

### Raw *sd* flow data
```{r}
#| label: flow_sdflowByRiverRaw1
#| fig-cap: "Standard deviation of stream flow (m3/s) for the West Brook against stream flow by river"
#| 
ggplot(cd1, aes(sdFlowByArea_flowExt, sdFlowByRiver)) +
  geom_point() +
  geom_abline(slope = 1, intercept = 0) +
  facet_grid(river~season, scales = 'free', labeller = global_labellerRiverSeasonInt_WB)

```

#### Frequency plots of sd flow data types

```{r}
#| label: flow_sdflowSDByRiverRawFreqs3
#| fig-cap: "Standard deviation in stream flow distributions, type (red = flow, blue = flowByRiver, green = flowByArea_flowExt)"

ggplot(cd1, aes(sdFlow)) +
  #geom_density(color = viridis(10)[1], adjust = 2.5) +
  geom_density(color = "red3", adjust = 2.5) +
  geom_density(aes(sdFlowByRiver), color = "blue3", adjust = 2.5) +
  geom_density(aes(sdFlowByArea_flowExt), color = "green4", adjust = 2.5) +
  geom_density(aes(sdFlowByArea_ByRiver), color = "orange2", adjust = 2.5) +
  labs(x = "Flow m3/s") +
  facet_grid(river~season, scales = 'free', labeller = global_labellerRiverSeasonInt_WB)
```

#### Time series plots of standard deviation of flow data types

```{r}
#| label: flow_sdflowByRiverRawTimeSeries
#| fig-cap: "Stream flow time series, flow type (blue = flowByRiver, green = flowByArea_flowExt)"

ggplot(cd1) +
  #geom_point(aes(date, meanFlowByRiver), color = "blue3") +
  #geom_line(aes(date, sdFlow), color = "red3") +
  geom_line(aes(date, sdFlowByRiver), color = "blue3") +
  geom_line(aes(date, sdFlowByArea_flowExt), color = "green4") +
  labs(x = "Date", y = "Stream flow") +
  facet_grid(river~season, scales = 'free', labeller = global_labellerRiverSeasonInt_WB)
```



## Scaled mean flow data
### Scaled by season

```{r}
#| label: flow_flowByRiverBySeason

ggplot(cd1, aes(meanFlowByArea_flowExtScaledBySeason, meanFlowByRiverScaledBySeason, color = cohort)) +
  geom_point() +
  geom_abline(slope = 1, intercept = 0) +
  facet_grid(river~season, scales = 'free', labeller = global_labellerRiverSeasonInt_WB)

ggplot(cd1, aes(meanFlowScaledBySeason)) +
  geom_density(color = "red3", adjust = 2.5) +
  geom_density(aes(meanFlowByRiverScaledBySeason), color = "blue3", adjust = 2.5) +
  geom_density(aes(meanFlowByArea_flowExtScaledBySeason), color = "green4", adjust = 2.5) +
  labs(x = "Flow type (red = flow, blue = flowByRiver, green = flowByArea_flowExt)") +
  facet_grid(river~season, scales = 'free', labeller = global_labellerRiverSeasonInt_WB)

ggplot(cd1, aes(meanTemperatureScaledBySeason)) +
  geom_density(color = "red3", adjust = 2.5) +
  #geom_density(aes(meanTemperatureByRiverScaledBySeason), color = "blue3", adjust = 2.5) +
  facet_grid(river~season, scales = 'free', labeller = global_labellerRiverSeasonInt_WB)
```

#### Time series plots of flow data types scaled by season

```{r}
#| label: flow_flowByRiverRawTimeSeriesScaledbyS
#| fig-cap: "Stream flow time series, flow type (blue = flowByRiver, green = flowByArea_flowExt)"

ggplot(cd1) +
  #geom_point(aes(date, meanFlowByRiver), color = "blue3") +
  geom_line(aes(date, meanFlowByRiverScaledBySeason), color = "blue3") +
  geom_line(aes(date, meanFlowByArea_flowExtScaledBySeason), color = "green4", adjust = 2.5) +
  labs(x = "Date", y = "Stream flow") +
  facet_grid(river~season, scales = 'free', labeller = global_labellerRiverSeasonInt_WB)
```

### Scaled by season and river

```{r}
#| label: flow_flowByRiverBySeasonRiver

ggplot(cd1, aes(meanFlowScaledBySeasonRiver, meanFlowByRiverScaledBySeasonRiver, color = cohort)) +
  geom_point() +
  geom_abline(slope = 1, intercept = 0) +
  facet_grid(river~season, scales = 'free', labeller = global_labellerRiverSeasonInt_WB)

ggplot(cd1, aes(meanFlowScaledBySeasonRiver)) +
  geom_density(color = "red3", adjust = 2.5, linewidth = 2) +
  geom_density(aes(meanFlowByRiverScaledBySeasonRiver), color = "blue3", adjust = 2.5, linewidth = 1) +
  geom_density(aes(meanFlowByArea_flowExtScaledBySeasonRiver), color = "green4", adjust = 2.5, linewidth = 0.75) +
  labs(x = "Flow type (red = flow, blue = flowByRiver, green = flowByArea_flowExt)") +
  facet_grid(river~season, labeller = global_labellerRiverSeasonInt_WB)#, scales = 'free')

```

#### Time series plots of flow data types scaled by season and river

```{r}
#| label: flow_flowByRiverRawTimeSeriesScaledbySR
#| fig-cap: "Stream flow time series, flow type (blue = flowByRiver, green = flowByArea_flowExt)"

ggplot(cd1) +
  #geom_point(aes(date, meanFlowByRiver), color = "blue3") +
  geom_line(aes(date, meanFlowByRiverScaledBySeasonRiver), color = "blue3") +
  geom_line(aes(date, meanFlowByArea_flowExtScaledBySeasonRiver), color = "green4", adjust = 2.5) +
  labs(x = "Date", y = "Stream flow") +
  facet_grid(river~season, scales = 'free', labeller = global_labellerRiverSeasonInt_WB)
```

Check cumulative flow\
Stream power as flow metric??  
  