## Get antenna data

```{r}
#| label: dataAntennaOptions
#| include: false
knitr::opts_chunk$set(message = FALSE, warning = FALSE)
```

```{r}
#| label: librariesdataAntenna
#| echo: false
#library(devtools)
#options(download.file.method = "wininet") or turn off VPN
#devtools::install_github('Conte-Ecology/westBrookData/getWBData')
library(tidyverse)
library(lubridate)
library(getWBData)
library(validate)
library(targets)


# devtools::install_github('bletcher/getPrepareWBData')
# use this if command above doesn't work: options(download.file.method = "wininet")
# library(getPrepareWBData)
```

## Raw data
Data are generated in `R/getDataAntenna-targets.R` [file](https://github.com/bletcher/wbBook_quarto_targets/blob/main/R/getDataAntenna_targets.R) 
Raw antenna data from the WB database (see [get data info chapter](https://bletcher.github.io/wbBook_quarto_targets/getData.html) for details) include these data frames:  

### Deployment details
```{r}
deploy <- tar_read(antennaDeployment_target)
str(deploy)
#write.csv(deploy, file = "./data/outForDownload/antenna/deploy.csv")
#
table(deploy$riverOrdered, deploy$riverMeter)
```

### Antenna log
```{r}
log <- tar_read(antennaLog_target)
str(log)
#write.csv(log, file = "./data/outForDownload/antenna/log.csv")

table(log$riverMeter,log$reader_id)
```


### Antenna status
Only dates
```{r}
status <- tar_read(antennaStatus_target)
str(status)
```

### Sites
Only dates
```{r}
sites <- tar_read(sites_target)
str(sites)
```


### Get antenna data observations target

```{r}
#| label: dataAntenna
#| cache: false
library(targets)
cdWB_antenna0 <- tar_read(cdWB_antenna0_target)
str(cdWB_antenna0)

# Save csv for downloading
# write.csv(cdWB_antenna0, file = "C:/Users/bletcher/OneDrive - DOI/projects/wbBook_quarto_targets/data/outForDownload/cdWB_antenna0.csv")
# Save rData file for downloading
# save(cdWB_antenna0, file = "C:/Users/bletcher/OneDrive - DOI/projects/wbBook_quarto_targets/data/outForDownload/cdWB_antenna0.RData")
table(cdWB_antenna0$riverMeter, cdWB_antenna0$riverOrdered, cdWB_antenna0$survey)
```

### Update antenna names  
Todd created a file `antennaLocationsHistory_correctionsAntennaNaming.xlsx` that has the updated names for the antennas. Names follow river_rkm format.  Merge in updated names below.  
```{r}
#| label: updateAntennaNames

antennaNames <- read.csv("./data/in/antennaLocationsHistory_correctionsAntennaNaming.csv") |>
  # remove left multiplexor anntenna names for now
  filter(!(antennaName %in% c(
    "WB_A1_RL_4259.55",
    #"WB_A1_RR_4259.55",
    "WB_A2_RL_4263.95"
    #"WB_A2_RR_4263.95"
  )))
antennaNames

cdWB_antenna <- cdWB_antenna0 |>
  left_join(antennaNames) |>
  dplyr::select(-survey, -sex, -river, -drainage)

# Save csv for downloading
# write.csv(cdWB_antenna, file = "C:/Users/bletcher/OneDrive - DOI/projects/wbBook_quarto_targets/data/outForDownload/cdWB_antenna.csv")
# Save rData file for downloading
# save(cdWB_antenna, file = "C:/Users/bletcher/OneDrive - DOI/projects/wbBook_quarto_targets/data/outForDownload/cdWB_antenna.RData")
table(cdWB_antenna$riverMeter, cdWB_antenna$riverOrdered)
```

