
## Sawmill, Shorey, Catamaran data
```{r}
#| label: getElectroNonWBDataOptionsDataRelease
#| include: false
knitr::opts_chunk$set(message = FALSE, warning = FALSE)
```

```{r}
#| label: librariesGetElectroDataDataRelease
#| echo: false
#library(devtools)
#options(download.file.method = "wininet") or turn off VPN
#devtools::install_github('Conte-Ecology/westBrookData/getWBData')
library(getWBData)
library(tidyverse)
library(lubridate)
library(targets)

# devtools::install_github('bletcher/getPrepareWBData')
# use this if command above doesn't work: options(download.file.method = "wininet")
#library(getPrepareWBData)
```

Data are generated in `R/getDataAll_noWB_targets.R` [file](https://github.com/bletcher/wbBook_quarto_targets/blob/main/R/getDataAll_noWB_targets.R)  

## Get data target  
Select variables for data release
```{r}
#| label: getNoWBDataDataRelease
#| cache: false

library(targets)
cdNoWB_all0 <- tar_read(cdNoWB_all_target) |>
  mutate(
    date = mdy(date),
    medianSampleDate = mdy(medianSampleDate)
  )

#cdNoWB_all0[2,]

str(cdNoWB_all0)

```

## Data checks  

```{r}
#| label: getNoWBDataSummary

table(cdNoWB_all0$drainage, cdNoWB_all0$river)

countTags <- cdNoWB_all0 |>
  group_by(tag, survey) |>
  summarize(count = n(), river = unique(riverTagged))
  
ggplot(countTags |> filter(survey %in% c("ANTENNA","NIGHT SEINE", "SHOCK", "SMOLT TRAP" ), river != ""), aes(count, fill = survey)) +
  geom_histogram() +
  scale_x_log10("Count per individual") +
  facet_grid(survey~river, scales = "free")
```

## Data checks  

```{r}
#| label: getNoWBDatLW_Summary
#| 
ggplot(cdNoWB_all0, aes(observedLength, observedMass)) +
  geom_point() +
  facet_wrap(~river)

cdNoWB_all <- cdNoWB_all0 |> 
  filter(!(observedLength < 100 & observedMass > 25)) |>
  filter(!(observedLength > 150 & observedMass < 10))

ggplot(cdNoWB_all, aes(observedLength, observedMass)) +
  geom_point() +
  facet_wrap(~river)

str(cdNoWB_all)
```

Date range
```{r}
#| label: dataRange
#|
min(cdNoWB_all$date)
max(cdNoWB_all$date)
```


## Summary data
```{r}
#| label: summaryTables
#| 
table(cdNoWB_all$river, cdNoWB_all$survey)

table(cdNoWB_all$river, cdNoWB_all$species)

table(cdNoWB_all$river, cdNoWB_all$season)

table(cdNoWB_all$river, cdNoWB_all$sampleNumber)

table(cdNoWB_all$river, cdNoWB_all$area)

```

Antenna sections from Shorey are missing from this that are text (A3, A4, SECTION -6 ANTENNA, TEMPORARY ANTENNA)
```{r}
#| label: summaryTables2
table(cdNoWB_all$river, as.numeric(cdNoWB_all$section))
```

Some formatting/filtering
```{r}
#| label: moreFiltering

cdNoWB_all <- cdNoWB_all |>
  mutate(
    source = tolower(source),
    source = ifelse(source == "sawmill river", "sawmill", source),
    sampleType = tolower(sampleType),
    survey = tolower(survey),
    area = tolower(area),
    estimatedYearOfStocking = tolower(estimatedYearOfStocking),
    aliveDead = tolower(aliveDead)
  )

```

## Save files for outputting
```{r}
# Save csv for downloading
# write.csv(cdNoWB_all, file = "C:/Users/bletcher/OneDrive - DOI/projects/wbBook_quarto_targets/data/out/dataRelease/cdNoWB_all_DR.csv")

# Save rData file for downloading
# save(cdNoWB_all, file = "C:/Users/bletcher/OneDrive - DOI/projects/wbBook_quarto_targets/data/outForDownload/cdNoWB_all.RData")
# 
```


## Metadata

#### Column explanations
__id__ Unique row ID from the original datasets.  

__source__ `shorey` = non-antenna data from shorey, `antenna tags` = antenna data from shorey,  `catamaran` = all data from catamaran, `sawmill` = all data from sawmill  

__sampleType__ `shorey` = non-antenna data from shorey, `stationary 1001` = antenna data from shorey, `stationary 2001` = antenna data from shorey, `catamaran` = all data from catamaran, `shorey`. The field is blank for sawmill.    

__drainage__ `catamaran` = Catamaran Brook drainage, `sawmill` = Sawmill River drainage, `shorey` = Shorey Brook drainage.  

__species__  
`bkt` = brook trout (*Salvelinus fontinalis*, native, self-sustaining population)  
`bnt` = brown trout (*Salmo trutta*, non-native, self-sustaining population)  
`ats` = Atlantic salmon (*Salmo salar*, stocked as fry in sawmill and shorey, natural reproduction in catamaran)
`splake` = splake (*Salvelinus namaycush X Salvelinus fontinalis*, hybrid of brook and lake trout - only in shorey)  

__river__ Same as __drainage__.    

__survey__ `antenna` = antenna data from shorey, `cc` = x from shorey `lf` = x from shorey, `night seine` = night seining from sawmill and shorey, `shock` = electroshocking sample, `smolt trap` = smolt trap data for salmon from catamaran and shorey.    

__tag__ PIT tag number, unique identifier, character. Tags are 12-mm, full duplex.    

__firstSeen__ Binary variable for whether the occasion is the first observation (1) or the fish, or not (0).  
 
__lastSeen__ Binary variable for whether the occasion is the last observation (1) or the fish, or not (0).   

__sampleNumber__ An identifier for the consecutive sample number for each river.  

__section__ Identifier for the 20m section where the fish was captured. This is ordered from downstream to upstream starting at 1 for the bottom of the main study area.  

__area__ `inside` = primary study area for each river. `above` and `above above` are sections above the main study area, `below` and `below below` are sections below the main study area, `antenna` are antenna sections  , `smolt trap` and `smolt trap new` represent the  smolt trap areas.    

__sampleName__ Represents the sample name for the sample and is a combination of "fish age"_"season number"_"season name".  

__season__ `Fall`, `Fall2` = second fall sample in shorey, `July` = July in shorey, `PostSmolt` = after the smolt sample, `PreSmolt` = before the smolt sample, `PreWinter` = before winter, `Smolt` = the smolt trap sample, `Summer`.  

__everEmMain__ Binary variable for whether the fish was ever observed in the main study area (1) or not (0).  

__riverTagged__ The river in which the fish was tagged.  

__areaTagged__ The area in which the fish was tagged.  

__riverKM__ River kilometer from the bottom of the study river.  

__date__ Date of observation. POSIXct format.  

__medianSampleDate__ Median date of the sampling occasion. POSIXct format.  

__fishNumber__ The consecutive fish number for each section and sample during sampling.  

__recap__ Binary variable for whether the fish was recaptured ("Y") or not ("0").  

__yearOfStocking__ Year of stocking when stocking occurred.  

__estimatedYearOfStocking__ Whether the year of stocking was estimated from fish length or not.   

__age__ Age of the fish in years.  

__observedLength__ Fish fork length in mm (+/-1 mm).  

__observedLength2__ Observed length lagged forward one observation for consecutive capture occasions. NA if not observed.  

__grLength__ Growth rate in length. (observedLength2 - observedLength) / sampleInterval (days).   

__grOstrovskyLength__ Growth rate in length after accounting for allometry. Based on Ostrovsky 1995.    

__observedMass__ Fish wet mass in g (+/- 0.1 g).  

__predMass__ Whether mass was predicted from length ("Yes") or measured ("No").  

__observedMass2__ Observed mass lagged forward one observation for consecutive capture occasions. NA if not observed.

__grMass__ 'Instantaneous' growth rate in mass. ln(lagObservedMass - observedMass) / sampleInterval.  

__grOstrovskyMass__ Growth rate in length after accounting for allometry. Based on Ostrovsky 1995.    

__grInterval__ Number of days between consecutive capture occasions.  

__cf__ Condition factor. (observedMass / observedMass^3) * 100.  

__mature__ Whether the fish was mature ("P") or not ("NP"). "P" stands for "precocious".  

__everMat__ Whether the fish was ever mature at any observation ("P") or not ("empty").  

__geneticSample__ Whether the fish was sampled for genetic analysis ("Y") or not ("N").  

__readerID__ PIT antenna reader ID for shorey data.  

__aliveDead__ `alive-moved` = returned alive to the stream in a new location, `alive-sample` = returned alive to the stream, `alive-sampled` = returned alive to the stream, `alive-samplemort` = mortality during workup.  

__antInstance__ PIT antenna instance number for individual fish.  

__pass__ Electrofishing pass 1, 2 or 3, 0 for antennas.  

__effort__ Electrofishing effort in seconds for each pass.  

__comment__ Comments from the field crew during sampling.  

__quarterWidth__ Width of the quarter in meters.  

__quarterLength__ Length of the quarter in meters.  





