# Introduction {.unnumbered}

**NOTE**: This is an internal document for model development. It is not intended for distribution and has not been approved by the USGS.


## The West Brook stream network

The Ecology Section at the USGS Conte Laboratory has studied fish in the West Brook (MA, USA) since 1997. The goal is to understand the strength and direction of drivers on fish growth, movement, reproduction and survival in a small stream network. We hope to provide a comprehensive understanding of fish population dynamics and ultimately individual fitness (natural selection and evolution) in the study area.

See below or [here](https://pitdata.ecosheds.org/) for interactive applications to explore the raw data.

```{r, echo = FALSE}
knitr::include_url("https://pitdata.ecosheds.org/", height = "850px")
```

Data extraction and manipulation and models are run using targets from the R package `targets`.
Run the following code to rebuild all the targets.  
`library(targets) library(quarto)`  
`tar_make()`   
Run this code to rebuild the book.  
`quarto::quarto_render(output_format = "html")`


## Mainstem and tributary names

There are four locations within the study area. Here are the primary names and aliases:

| Primary name   | Alias | Alias-long     | River # |
|----------------|-------|----------------|---------|
| West Brook     | WB    | West Brook     | 1       |
| Jimmy Brook    | OL    | Open Large     | 2       |
| Mitchell Brook | OS    | Open Small     | 3       |
| OBear Brook    | IL    | Isolated Large | 4       |

: River names and aliases {#tbl-rivers}

```{r}
#if/when .qmd's are in quarto subdir
#tar_config_set(store = "../_targets")
```

