# Production model Abundance model - river 4 (IS)

Simple JS model integrated with a growth in weight model to get phi, p, and growth estimates to develop a production model. The CJS model in modelProduction_CJS_OB.qmd was used to test out the combined mass/survival model. here here take from that .qmd page and run abundance models that start at ageInSamples 1 (or 0). So we can estimate biomass for all fish estimated to be alive for any occasion.  

In all the models below, 1 = not observed and 2 = observed in the input *encounter data*.\
Encounter data are available [here](https://drive.google.com/drive/folders/1UmPv49xL-mzUBndqjRISylNO3q3Zs-cm) in the `eh.csv` file. Weight data are in `weight.csv`

```{r}
#| label: librariesModelsNimbleProduction
#| echo: false
library(getWBData)
library(lubridate)
library(kableExtra)
library(GGally)
library(nimble)
library(nimbleEcology)
library(MCMCvis)
library(tidyverse)
library(targets)

library(shiny)
library(tibble)
```

```{r}
#| label: globalModelsNimbleProduction
#| include: false
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
```

TO DO:

1) Evaluate need to run full all-cohort models vs. cohort-specific models.\

2) Evaluate first or second order quadratic models (model 2 0r 3).  

3) We are going to build off of the OB models here by  
  - CJS -> JS in OB
  - OB JS model to WB by adding **species** and **movement* among rivers. Could run species models separately  
  
  


All target files are in `./R`. From `getDataElectro_targets.R` we have the following target files: cdWB_electro_target.  

Encounter history files come from `dataCMR_OB_2002_2014_target` or `dataCMR_OB_singleCohorts_target` in `getDataCMR_targets.R`  
`getEH_AIS()` is the main function to create the CMR list.  

Run the model with `run_production_OB_model()` which is in `./R/modelProduction_JS_OB_functionsToSource.R`



Model code is in `./models/production/modelProduction_JS_OB_functionsToSource.R`\
Model is run 'by hand' in `./models/modelProduction_JS_OB_makeFile.R`\
Functions for this qmd file in `./models/qmdProduction_JS_OB_functionsToSource.R`\

There are **3** models here:
All cohorts:  
  model **1**, with quadratic weight effect on phi
By cohort:  
  model **2**, with linear weight effect on phi
  model **3**, with quadratic weight effect on phi

### How many ageInSamples to include?

```{r}
#| label: production

all <- tar_read(cdWB_electro_target)
table(all %>% filter(river == "wb obear")|> dplyr::select(ageInSamples))
table(all %>% filter(river == "wb obear")|> dplyr::select(cohort,ageInSamples))

cohorts <- 2002:2014

```

## All Cohorts, modelNum **1**: phi(i,t,c) \* g(i,t,c), p(i,t) model


## By Cohort, modelNum **2**: phi(i,t) \* g(i,t), p(i,t) model


Using model #3 from `modelsCMR_Growth_NN_OB.qmd` as a staring point for the models, but adapting the model by\
Prob not 1) Extending AgeInSamples from 1-11 to 1-x to allow bigger fish to be present for the production estimates.  
Prob not 2) Loop over first\[i\]:lastAIS so fish have the chance to survive to large size.\


Model with phi and p for each age-in-samples (t = column in the encounter history file) and individual (i)

Probability of survival (phi) model structure:

```         
      logit(phi[i,t]) <- 
        phiInt[i,t] + 
        phiBeta[1,i,t] * weight[i,t] 
```

Probability of capture (p) model structure:

p(t,i) \<- pInt(t,i)

Growth rate (gr) model structure:

gr(t,i) \<- grInt(t)

Survival/growth rate interaction:

Multiplicative

Model runs:\

## By Cohort, modelNum **3**: phi(i,t) \* g(i,t)^2, p(i,t) model


Using model #3 from `modelsCMR_Growth_NN_OB.qmd` as a staring point for the models, but adapting the model by\
Prob not 1) Extending AgeInSamples from 1-11 to 1-x to allow bigger fish to be present for the production estimates.  
Prob not 2) Loop over first\[i\]:lastAIS so fish have the chance to survive to large size.\


Model with phi and p for each age-in-samples (t = column in the encounter history file) and individual (i)

Probability of survival (phi) model structure:

```         
      logit(phi[i,t]) <- 
        phiInt[i,t] + 
        phiBeta[1,i,t] * weight[i,t] + 
        phiBeta[2,i,t] * weight[i,t] * weight[i,t] 
```

Probability of capture (p) model structure:

p(t,i) \<- pInt(t,i)

Growth rate (gr) model structure:

gr(t,i) \<- grInt(t)

Survival/growth rate interaction:

Multiplicative

Model runs:\

### Retrieve model results

```{r}
#| label: productionSource_3
#| cache: false
library(targets)
library(nimble)
# Following https://oliviergimenez.github.io/bayesian-cr-workshop/worksheets/4_demo.html
# 
# 

# To get the mMCMCSummaryRMNA funcion which I adapted to deal with NAs
#source('./models/production/modelProduction_JS_OB_functionsToSource.R')
#source('./models/production/qmdProduction_JS_OB_functionsToSource.R')

modelNum <- 3 # phi * growth
```


