---
editor: 
  markdown: 
    wrap: 72
---

# Growth in mass

```{r}
#| label: dataModelFlowOptions
#| include: false
knitr::opts_chunk$set(message = FALSE, warning = FALSE)
```

```{r}
#| label: librariesModelsFlow
#| echo: false

######################################################

# for gams
library(gratia)
library(arm)
library(relaimpo)
library(GGally)
# library(linkedModels)
library(arrayhelpers)
library(mgcv)
#####################################################

library(getWBData)
library(lubridate)
library(kableExtra)
library(GGally)
library(lme4)
library(targets)
library(relaimpo)
library(caret)
library(pscl)

# for xgboost models
library(rsample)      # data splitting 
library(gbm)          # basic implementation
library(xgboost)      # a faster implementation of gbm
#library(h2o)          # a java-based platform
library(pdp)          # model visualization
library(lime)
library(vtreat)
library(Ckmeans.1d.dp)

library(tidyverse)

`%notin%` <- Negate(`%in%`)
```

Distribution of growth, negative growth\
- Seasonal differences\
- Species differences in size/growth/CF - Patterns of negative growth

Causes of neg growth\
- Extrinsic\
- river, season, flow, temp\
- intrinsic\
- Sex (fall), life history, section, size, age, previous growth (CF?),
movement (section/river)\
consequences of neg growth - survival\
- future growth (length and mass)\
- movement (section/river)\
- reprod success?\
- local density (con and hetero)

Probs with growth in mass\
spawning loss mass in the fall, esp in OB

Interaction between flow:temp, diff btw WB (groundwater dominated at low
flow so cold) and tribs (combo of surface and gw?)

Compare growth with t/f cycle diagrams for each river

compare condition factor among species - looks like bkt growing faster
in weight

flow/temp circles with gr on top

## Get data

### Get electrofishing data

```{r}
#| label: readInAllData
#| cache: false
library(targets)
cd <- tar_read(cdWB_electro_target) # raw data
cd1 <- tar_read(cd1_target) # from modelGrowthInMass_target
propNegSRS <- tar_read(propNegSRS_target)
propNegSRsN <- tar_read(propNegSRsN_target)

indCountsBySpp <- tar_read(indCountsBySpp_target) 
indCounts <- tar_read(indCounts_target) 
propNegLabels <- tar_read(propNegLabels_target)

negGr_beforeLast <- tar_read(negGr_beforeLast_target)

meanNegSRsN <- tar_read(meanNegSRsN_target)
meanNegSRsNWide <- tar_read(meanNegSRsNWide_target)

# df for pairs plots
cd1Wide_grWeight <- tar_read(cd1Wide_grWeight_target)
cd1Wide_observedWeight <- tar_read(cd1Wide_observedWeight_target)

dGAM <- tar_read(dGAM_target) # from modelGrowthInMass_target
```

### Get environmental data

```{r}
#| label: readInAllDataEnv
#| cache: false
library(targets)
dummy=0 # change to force cache update??
envIn <- tar_read(envIn_target)
envIn_propNeg <- tar_read(envIn_propNeg_target)
```

## Examine raw data

### Limit data to consecutive sample captures (cd1)

Previous growth models with length used all possible observations for a
fish and interpolated missing observations. Here, we are just looking at
consecutive captures to estimate growth over just that interval.

```{r}
#| label: consecSamples
# table(as.numeric(cd$sampleNumber), as.numeric(cd$lagSampleNumber), cd$season, cd$year) |> 
#   data.frame() |> 
#   filter(Freq > 0) |> 
#   arrange(Var1, Var2) |> 
#   rename(firstCapt = Var1, secondCapt = Var2, season = Var3, year = Var4) |> 
#   filter(year == 2010)


##### in targets
# cd1 <- cd |> 
#   filter(sampleNumberDiff == 1,
#          tag %notin% c('1bf20ff490', '1bf20ebe4e')) |> 
#   mutate(negGrowth = grWeight < 0,
#          month = month(date),
#          dummy = 1) # change to force cache update
```

### Find outlier growth rates

#### Negative outlier growth

`1bf20ebe4e` and `1bf20ff490` have incorrect weights leading to very
negative growth rates.\
Exclude these tags in `cd1_target`. The rest of the quite negative
growth fish seem plausible (remember we can't account for stomach
contents or egg loss).

```{r}
#| label: negOutlier
cd |> 
  arrange(grWeightS) |> 
  dplyr::select(tag, species, season, sampleNumber, lagSampleNumber, observedWeight, lagObservedWeight, grWeightS)
```

#### Positive outlier growth

Lots of very fast growth mostly in the spring. No clear way to exclude
fish. `1c2c582218` is excluded in `cd1_target` because of very high
growth in length.

```{r}
#| label: posOutlier
cd1 |> 
  arrange(desc(grWeightS)) |> 
  dplyr::select(tag, species, season, sampleNumber, lagSampleNumber, observedWeight, lagObservedWeight, grWeightS) |> 
  print(n = 10)
```

### Raw length, weight data graphs

#### Distribution of growth rates

```{r}
#| label: rawlengthWeight
ggplot(cd1, aes(grWeight, color = speciesGG)) +
  geom_freqpoly(bins = 50) +
  labs(x = "Growth in mass (/d)", y = "Frequency") +
  scale_color_discrete() +
  facet_wrap(~seasonGG)

ggplot(cd1, aes(grLength, color = speciesGG)) +
  geom_freqpoly(bins = 50) +
  labs(x = "Growth in Length (mm/d)", y = "Frequency") +
  scale_color_discrete() +
  facet_wrap(~seasonGG)
```

### Length vs weight by species

Very few outliers

```{r}
#| label: len_v_weight
plotLenW <- function(d, s){
  ggplot(d |> filter(species == s), aes(observedLength, observedWeight)) +
    geom_point(alpha = 0.5, size = 1) +
    scale_x_log10() +
    scale_y_log10() +
    #ggtitle(s) +
    labs(x = "Length (mm)", y = "Mass (mg)") +
    scale_color_discrete() +
    facet_grid(riverGG~year, scales = 'free')
}
```

::: panel-tabset
#### Brook trout

```{r}
#| label: len_v_weight_bkt

plotLenW(cd1, "bkt")
```

#### Brown trout

```{r}
#| label: len_v_weight_bnt

plotLenW(cd1, "bnt")
```

#### Atlantic salmon

```{r}
#| label: len_v_weight_ats

plotLenW(cd1, "ats")
```
:::

### Weight vs weight at next sample

```{r}
#| label: weight_v_weight
plotWW <- function(d, spp){
  ggplot(d |> filter(species == spp), aes(observedWeight, lagObservedWeight, color = seasonGG)) +
    geom_point(alpha = 0.15) +
    guides(color = guide_legend(override.aes = list(alpha = 1))) +
    labs(x = "Mass (mg)", y = "Mass (mg) at next sample") +
    scale_color_discrete() +
    facet_wrap(~riverGG, scales = 'free')
}
```

::: panel-tabset
#### Brook trout

```{r}
#| label: weight_v_weight_bkt

plotWW(cd1, "bkt")
```

#### Brown trout

```{r}
#| label: weight_v_weight_bnt

plotWW(cd1, "bkt")
```

#### Atlantic salmon

```{r}
#| label: weight_v_weight_ats

plotWW(cd1, "ats")
```
:::

### Growth rate pairs plots

By species, all cohorts\
Row and column labels in the pairs plots are 'age'\_'season'

```{r}
#| label: pairs

plotPairs <- function(d0, sppIn, cohortIn, firstSamp = 1, lastSamp = 9, upperBlank = FALSE){
  d <- d0 |> 
    filter(speciesGG == sppIn, cohort %in% cohortIn) |> 
    select_if(~any(!is.na(.))) |>  # remove all-NA columns
    filter(firstObserved != lastObserved - 1) # remove single obs fish
  
  lowerFn <- function(data, mapping, method = "lm", ...) {
    ggplot(data = data, mapping = mapping) +
      geom_vline(xintercept = 0, color = 'darkgrey') +
      geom_hline(yintercept = 0, color = 'darkgrey') +
      geom_abline(intercept = 0, slope = 1) +
      geom_smooth(aes(color = factor(cohort)), method = method, se = FALSE, ...) +
      geom_point(aes(color = factor(cohort))) +
      geom_smooth(color = 'black', size = 1.1, method = method, se = FALSE, ...) 
  }

  if(upperBlank) {
    out <- ggpairs(
      d |> 
        dplyr::select(-c(speciesGG, tag, firstObserved, lastObserved)),
      columns = (firstSamp+1):(lastSamp+1), # +1 to skip cohort column
      lower = list(continuous = wrap(lowerFn, method = "lm")),
      upper = "blank"
    ) 
  } else {
    out <- ggpairs(
      d |> 
        dplyr::select(-c(speciesGG, tag, firstObserved, lastObserved)),
      columns = (firstSamp+1):(lastSamp+1), # +1 to skip cohort column
      lower = list(continuous = wrap(lowerFn, method = "lm"))
    )
  }
  return(out)
}

```

::: panel-tabset
#### Brook trout

```{r}
#| label: pairs_bkt

plotPairs(cd1Wide_grWeight, "Brook trout", 2000:2015, 1, 9)
#plotPairs(cd1Wide_observedWeight, "Brook trout", 2000:2015)
```

#### Brown trout

```{r}
#| label: pairs_bnt

plotPairs(cd1Wide_grWeight, "Brown trout", 2000:2015, 1, 9)
#plotPairs(cd1Wide_observedWeight, "Brook trout", 2000:2015)
```

#### Atlantic salmon

```{r}
#| label: pairs_ats

plotPairs(cd1Wide_grWeight, "Atlantic salmon", 2000:2015, 1, 8) # missing data in column 9
#plotPairs(cd1Wide_observedWeight, "Atlantic salmon", 2000:2015)
```
:::

#### Brook trout by cohort

Row and column labels are 'age'\_'season'\
Color by river - make new function

::: panel-tabset
#### 2000

```{r}
#| label: pairs_bkt2000

plotPairs(cd1Wide_grWeight, "Brook trout", 2000, 1, 8, upperBlank = TRUE)
```

#### 2001

```{r}
#| label: pairs_bkt2001

plotPairs(cd1Wide_grWeight, "Brook trout", 2001, upperBlank = TRUE)
```

#### 2002

```{r}
#| label: pairs_bkt2002

plotPairs(cd1Wide_grWeight, "Brook trout", 2002, 1, 9, upperBlank = TRUE)
```

#### 2003

```{r}
#| label: pairs_bkt2003

plotPairs(cd1Wide_grWeight, "Brook trout", 2003, upperBlank = TRUE)
```

#### 2004

```{r}
#| label: pairs_bkt2004

plotPairs(cd1Wide_grWeight, "Brook trout", 2004, upperBlank = TRUE)
```

#### 2005

```{r}
#| label: pairs_bkt2005

plotPairs(cd1Wide_grWeight, "Brook trout", 2005, upperBlank = TRUE)
```

#### 2006

```{r}
#| label: pairs_bkt2006

plotPairs(cd1Wide_grWeight, "Brook trout", 2006, upperBlank = TRUE)
```

#### 2007

```{r}
#| label: pairs_bkt2007

plotPairs(cd1Wide_grWeight, "Brook trout", 2007, upperBlank = TRUE)
```

#### 2008

```{r}
#| label: pairs_bkt2008

plotPairs(cd1Wide_grWeight, "Brook trout", 2008, upperBlank = TRUE)
```

#### 2009

```{r}
#| label: pairs_bkt2009

plotPairs(cd1Wide_grWeight, "Brook trout", 2009, upperBlank = TRUE)
```

#### 2010

```{r}
#| label: pairs_bkt2010

plotPairs(cd1Wide_grWeight, "Brook trout", 2010, 1, 8, upperBlank = TRUE)
```

#### 2011

```{r}
#| label: pairs_bkt2011

plotPairs(cd1Wide_grWeight, "Brook trout", 2011, 1, 7, upperBlank = TRUE)
```

#### 2012

```{r}
#| label: pairs_bkt2012

plotPairs(cd1Wide_grWeight, "Brook trout", 2012, upperBlank = TRUE)
```

#### 2013

```{r}
#| label: pairs_bkt2013

plotPairs(cd1Wide_grWeight, "Brook trout", 2013, upperBlank = TRUE)
```

#### 2014

```{r}
#| label: pairs_bkt2014

plotPairs(cd1Wide_grWeight, "Brook trout", 2014, 1, 5, upperBlank = TRUE)
```
:::

#### Consecutive season growth comparisons by age_season

Brook trout

```{r}
#| label: pairsByAgeSeason

ggplot(cd1Wide_grWeight |> filter(speciesGG == "Brook trout", cohort %in% 2000:2014, `0_4` < 0.01), aes(`0_3`, `0_4`, color = factor(cohort))) +
  geom_point() +
  geom_vline(xintercept = 0, color = 'darkgrey') +
  geom_hline(yintercept = 0, color = 'darkgrey') +
  #geom_abline(intercept = 0, slope = 1) +
  geom_smooth(method = "lm", se = FALSE, color = '#45454d') +
  guides(colour = FALSE) +
  facet_wrap(~cohort)

ggplot(cd1Wide_grWeight |> filter(speciesGG == "Brook trout", cohort %in% 2000:2014), aes(`0_4`, `1_1`, color = factor(cohort))) +
  geom_point() +
  geom_vline(xintercept = 0, color = 'darkgrey') +
  geom_hline(yintercept = 0, color = 'darkgrey') +
  #geom_abline(intercept = 0, slope = 1) +
  geom_smooth(method = "lm", se = FALSE, color = '#45454d') +
  guides(colour = FALSE) +
  facet_wrap(~cohort)

ggplot(cd1Wide_grWeight |> filter(speciesGG == "Brook trout", cohort %in% 2000:2014), aes(`1_1`, `1_2`, color = factor(cohort))) +
  geom_point() +
  geom_vline(xintercept = 0, color = 'darkgrey') +
  geom_hline(yintercept = 0, color = 'darkgrey') +
  #geom_abline(intercept = 0, slope = 1) +
  geom_smooth(method = "lm", se = FALSE) +
  guides(colour = FALSE) +
  facet_wrap(~cohort)

ggplot(cd1Wide_grWeight |> filter(speciesGG == "Brook trout", cohort %in% 2000:2014), aes(`1_2`, `1_3`, color = factor(cohort))) +
  geom_point() +
  geom_vline(xintercept = 0, color = 'darkgrey') +
  geom_hline(yintercept = 0, color = 'darkgrey') +
  #geom_abline(intercept = 0, slope = 1) +
  geom_smooth(method = "lm", se = FALSE, color = '#45454d') +
  guides(colour = FALSE) +
  facet_wrap(~cohort)
 
ggplot(cd1Wide_grWeight |> filter(speciesGG == "Brook trout", cohort %in% 2000:2012), aes(`1_1`, `2_1`, color = factor(cohort))) +
  geom_point() +
  geom_vline(xintercept = 0, color = 'darkgrey') +
  geom_hline(yintercept = 0, color = 'darkgrey') +
  geom_abline(intercept = 0, slope = 1) +
  geom_smooth(method = "lm", se = FALSE, color = '#45454d') +
  guides(colour = FALSE) +
  facet_wrap(~cohort)
 
```

### Growth rate in length vs growth rate in mass

```{r}
#| label: grLen_v_grWeight
ggplot(cd1, aes(grLength, grWeight)) +
  geom_point(aes(size = observedWeight), alpha = 0.02) +
  guides(size = guide_legend(override.aes = list(alpha = 1))) +
  labs(x = "Growth in length", y = "Growth in mass") +
  scale_color_discrete() +
  facet_grid(seasonGG~speciesGG)

plotGrLenMass <- function(d, s){
  ggplot(cd1 |> filter(species==s), aes(grLength, grWeight, color = factor(year))) +
    geom_point(alpha = 0.5) +
    geom_smooth(method="lm", se=FALSE) +
    guides(size = guide_legend(override.aes = list(alpha = 1))) +
    labs(x = "Growth in length", y = "Growth in mass") +
    scale_color_discrete() +
    facet_wrap(~seasonGG, scales = 'free')  
}
#  facet_wrap(~year)
```

::: panel-tabset
#### Brook trout

```{r}
#| label: grLen_v_grWeight_bkt

plotGrLenMass(cd1, "bkt")
```

#### Brown trout

```{r}
#| label: grLen_v_grWeight_bnt

plotGrLenMass(cd1, "bnt")
```

#### Atlantic salmon

```{r}
#| label: grLen_v_grWeight_ats

plotGrLenMass(cd1, "ats")
```
:::

No relationship between means of neg and pos growth

```{r}
#| label: posNegGR
#| 

ggplot(meanNegSRsN |> filter(speciesGG == "Brook trout"), aes(year, meanNegPos, color = negGrowth, group = negGrowth)) +
  geom_point() +
  geom_line() +
  geom_hline(yintercept = 0) +
  facet_grid(seasonGG~riverGG)

ggplot(meanNegSRsNWide, aes(`FALSE`, `TRUE`)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  facet_grid(seasonGG~riverGG)
```

## Flow and temperature growth relationships

### Mean flow effect on growth?

```{r}
#| label: flow_weight

plotFlowGR <- function(d, r){
ggplot(d |> filter(river == r), aes(meanFlowByRiver, grWeight, color = speciesGG)) +
  geom_point(alpha = 0.05)  +
  geom_smooth(method = "lm", se = FALSE) +
  geom_smooth(se = FALSE) +
  geom_smooth(method = "lm", se =FALSE, formula = y~poly(x,2)) +
  labs(x = "Mean flow", y = "Growth in mass") +
  scale_color_discrete() +
  #ggtitle(r) +
  facet_grid(seasonGG~speciesGG)#, scales = 'free')
}

```

::: panel-tabset
#### West Brook

```{r}
#| label: flow_weight_wb

plotFlowGR(cd1, "west brook")
```

#### Open Large

```{r}
#| label: flow_weight_ol

plotFlowGR(cd1, "wb jimmy")
```

#### Open Small

```{r}
#| label: flow_weight_os

plotFlowGR(cd1, "wb mitchell")
```

#### Isolated small

```{r}
#| label: flow_weight_is

plotFlowGR(cd1, "wb obear")
```
:::

### Mean temperature effect on growth?

Maybe in Jimmy for bnt, or for ats

```{r}
#| label: temp_grWeight

plotTempGR <- function(d, s){
  ggplot(d |> filter(seasonGG == s), aes(meanTemperature, grWeight, color = speciesGG)) +
    geom_point(alpha = 0.5)  +
    geom_smooth(method = "lm", se = FALSE, color = 'darkgrey') +
    geom_smooth(se = FALSE, color = 'darkgrey') +
    geom_smooth(method = "lm", se =FALSE, formula = y~poly(x,2), color = 'darkgrey') +
    labs(x = "Mean temperature (c)", y = "Growth in mass") +
    scale_color_discrete() +
   # ggtitle(s) +
    facet_grid(speciesGG~riverGG)#, scales = 'free')
}

```

::: panel-tabset
#### Spring

```{r}
#| label: temp_grWeight_spring

plotTempGR(cd1, "Spring")
```

#### Summer

```{r}
#| label: temp_grWeight_summer

plotTempGR(cd1, "Summer")
```

#### Autumn

```{r}
#| label: temp_grWeight_autumn

plotTempGR(cd1, "Autumn")
```

#### Winter

```{r}
#| label: temp_grWeight_winter

plotTempGR(cd1, "Winter")
```
:::

## Proportion of fish with **negative growth**

### Proportion of fish with negative growth by species/river/season

```{r}
#| label: neg_pos
ggplot(cd1 |> filter(!is.na(negGrowth)), aes(negGrowth)) +
  geom_bar() +
  facet_grid(seasonGG~speciesGG+riverGG, scales = "free")
```

### Proportion of fish with negative growth by species/season

```{r}
#| label: propNegBySeason

ggplot(propNegSRS |> filter(n > 50), aes(seasonGG, propNeg, color = riverGG, group = riverGG)) +
  geom_point(aes(size = n)) +
  geom_line() +
  labs(x = "Season", y = "Proportion of fish with negative growth in mass") +
  scale_color_discrete() +
  facet_wrap(~speciesGG)
```

### Proportion of fish with negative growth by species/river/year

```{r}
#| label: propNegWB_BKT

plotPropNegByRiver <- function(d, spp){
  ggplot(d |> filter(n > 10,  speciesGG == spp), 
         aes(year, propNeg, color = factor(seasonGG))) +
    geom_point(aes(size = n)) +
    geom_line() +
    labs(x = "Year", y = "Proportion of fish with negative growth in mass") +
    scale_color_discrete() +
    facet_wrap(~riverGG)
}
```

::: panel-tabset
#### Brook trout

```{r}
#| label: propNegWB_BKT_bkt

plotPropNegByRiver(propNegSRsN, "Brook trout")
```

#### Brown trout

```{r}
#| label: propNegWB_BKT_bnt

plotPropNegByRiver(propNegSRsN |> filter(riverGG != "Open Small"), "Brown trout")
```

#### Atlantic salmon

```{r}
#| label: propNegWB_BKT_ats

plotPropNegByRiver(propNegSRsN, "Atlantic salmon")
```
:::

### Proportion of fish with negative growth by species/river/year

Break down season by year, compare across rivers

```{r}
#| label: neg_posN

ggplot(propNegSRsN |> filter(n > 10), aes(year, propNeg, color = riverGG)) +
  geom_point(aes(size = n)) +
  geom_line() +
  labs(x = "Year", y = "Proportion of fish with negative growth in mass") +
  scale_color_discrete() +
  facet_grid(seasonGG~speciesGG)

```

## Environmental effects on proportion of fish with negative growth

### Proportion of fish with negative growth by **flow**

```{r}
#| label: negGrowth_flow

ggplot(propNegSRsN |> filter(n > 10, riverGG == "West Brook"), aes(mF, propNeg, color = riverGG)) +
  geom_point(aes(size = n)) +
  geom_smooth(method = "lm", se = FALSE) +
  scale_y_continuous("Proportion of fish with negative growth in mass") +
  scale_x_log10("Mean flow") +
  scale_color_discrete() +
  ggtitle("West Brook") +
  facet_grid(seasonGG~speciesGG)

ggplot(propNegSRsN |> filter(n > 10, riverGG != "West Brook"), aes(mF, propNeg, color = riverGG)) +
  geom_point(aes(size = n)) +
  geom_smooth(method = "lm", se = FALSE) +
  scale_y_continuous("Proportion of fish with negative growth in mass") +
  scale_x_log10("Mean flow") +
  scale_color_discrete() +
  ggtitle("Tributaries") +
  facet_grid(seasonGG~speciesGG)
```

### Proportion of fish with negative growth by **temperature**

```{r}
#| label: negGrowth_temp

plotNegGrowth_temp <- function(d, s){
  ggplot(d |> filter(n > 10, seasonGG == s), aes(mT, propNeg, color = riverGG)) +
    geom_point(aes(size = n)) +
    geom_smooth(method = "lm", se = FALSE) +
    scale_y_continuous("Proportion of fish with negative growth in mass") +
    scale_x_continuous("Mean temperature (C)") +
    scale_color_discrete() +
    ggtitle(s) +
    facet_grid(riverGG~speciesGG)
}

```

::: panel-tabset
#### Spring

```{r}
#| label: negGrowth_tempSPring
plotNegGrowth_temp(propNegSRsN, "Spring")
```

#### Summer

```{r}
#| label: negGrowth_tempSummer
plotNegGrowth_temp(propNegSRsN, "Summer")
```

#### Autumn

```{r}
#| label: negGrowth_tempAutumn
plotNegGrowth_temp(propNegSRsN, "Autumn")
```

#### Winter

```{r}
#| label: negGrowth_tempWinter
plotNegGrowth_temp(propNegSRsN, "Winter")
```
:::

### Proportion of fish with negative growth by **3 species fish counts**

```{r}
#| label: negGrowth_counts3Spp

ggplot(propNegSRsN |> filter(n > 10, riverGG == "West Brook"), aes(nInd, propNeg, color = riverGG)) +
  geom_point(aes(size = n)) +
  geom_smooth(method = "lm", se = FALSE) +
  scale_y_continuous("Proportion of fish with negative growth in mass") +
  scale_x_log10("3 species count") +
  scale_color_discrete() +
  ggtitle("West Brook") +
  facet_grid(seasonGG~speciesGG)

ggplot(propNegSRsN |> filter(n > 10, riverGG != "West Brook"), aes(nInd, propNeg, color = riverGG)) +
  geom_point(aes(size = n)) +
  geom_smooth(method = "lm", se = FALSE) +
  scale_y_continuous("Proportion of fish with negative growth in mass") +
  scale_x_log10("3 species count") +
  scale_color_discrete() +
  ggtitle("Tributaries") +
  facet_grid(seasonGG~speciesGG)
```

### Proportion of fish with negative growth by **conspecific fish counts**

```{r}
#| label: negGrowth_counts1Spp

ggplot(propNegSRsN |> filter(n > 10, riverGG == "West Brook"), aes(nIndBySpp, propNeg, color = riverGG)) +
  geom_point(aes(size = n)) +
  geom_smooth(method = "lm", se = FALSE) +
  scale_y_continuous("Proportion of fish with negative growth in mass") +
  scale_x_log10("conspecific count") +
  scale_color_discrete() +
  ggtitle("West Brook") +
  facet_grid(seasonGG~speciesGG)

ggplot(propNegSRsN |> filter(n > 10, riverGG != "West Brook"), aes(nIndBySpp, propNeg, color = riverGG)) +
  geom_point(aes(size = n)) +
  geom_smooth(method = "lm", se = FALSE) +
  scale_y_continuous("Proportion of fish with negative growth in mass") +
  scale_x_log10("conspecific count") +
  scale_color_discrete() +
  ggtitle("Tributaries") +
  facet_grid(seasonGG~speciesGG)
```

## F/T cycle graphs

### Mean flow and temperature

```{r}
#| label: flow_temp_cycle_bySpecies
plot_mT_mF <- function(d, s, plotText){
  ggplot(d |> filter(n > 25, speciesGG == s), aes(mT, mF_log10, color = factor(seasonGG))) +
    geom_point(aes(size = propNeg)) +
    {if(plotText) geom_text(aes(label = year), vjust = 0, nudge_y = -0.075, check_overlap = FALSE, size = 3)} +
    labs(x = "Mean temperature (C)", y = "log10 of mean Flow") +
    scale_color_discrete() +
    facet_wrap(~riverGG, scales = "free")
}

```

::: panel-tabset
#### Brook trout

```{r}
#| label: flow_temp_cycle_bySpecies_bkt

plot_mT_mF(propNegSRsN, "Brook trout", FALSE)
```

#### Brown trout

```{r}
#| label: flow_temp_cycle_bySpecies_bnt

plot_mT_mF(propNegSRsN, "Brown trout", FALSE)
```

#### Atlantic salmon

```{r}
#| label: flow_temp_cycle_bySpecies_ats

plot_mT_mF(propNegSRsN, "Atlantic salmon", FALSE)
```
:::

### Mean flow and temperature with year labels

#### panel-tabset

### Brook trout

```{r}
#| label: flow_temp_cycle_bySpeciesYear_bkt

plot_mT_mF(propNegSRsN, "Brook trout", TRUE)
```

#### Brown trout

```{r}
#| label: flow_temp_cycle_bySpeciesYear_bnt

plot_mT_mF(propNegSRsN, "Brown trout", TRUE)
```

#### Atlantic salmon

```{r}
#| label: flow_temp_cycle_bySpeciesYear_ats

plot_mT_mF(propNegSRsN, "Atlantic salmon", TRUE)
```

:::

Mean flow and temperature values for each season and year for the West
Brook\
Not very informative

```{r}
#| label: flow_temp_cycle_bySpeciesYear2
#| 
ggplot(propNegSRsN |> filter(n > 10, riverGG == 'West Brook') |> arrange(riverGG, speciesGG, year, seasonGG), aes(mT, mF_log10)) +
  geom_line() +
  geom_point(aes(color = factor(seasonGG))) +
  facet_wrap(~year)
```

### Yearly temperature/flow cycles of daily data

```{r}
#| label: mergePropNegPlot

plotMerged <- function(d, r, minYear = 2000, maxYear = 2017){
  ggplot(d |> filter(riverGG == r, year %in% c(minYear:maxYear)), 
         aes(temperature, log10(flowByRiverm3s), color = seasonGG, group = seasonGG)) +
    geom_point() +
    geom_smooth(se = FALSE) +
    #geom_line() +
    labs(x = "Temperature (C)", y = "log10 of Flow") +
    scale_color_discrete() +
    ggtitle(r) +
    facet_wrap(~year, scales = 'free')
}
# this plots f/t for all three species - good for filling in gaps

```

::: panel-tabset
#### West Brook

```{r}
#| label: mergePropNegPlot_wb

plotMerged(envIn_propNeg, "West Brook", 2000, 2017)
```

#### Open Large

```{r}
#| label: mergePropNegPlot_ol

plotMerged(envIn_propNeg, "Open Large", 2002, 2017)
```

#### Open Small

```{r}
#| label: mergePropNegPlot_os

plotMerged(envIn_propNeg, "Open Small", 2002, 2017)
```

#### Isolated small

```{r}
#| label: fmergePropNegPlot_is

plotMerged(envIn_propNeg, "Isolated Small", 2002, 2017)
```
:::

### Smooths of temperature/flow cycles of daily data by season

```{r}
#| label: plotMergedBySeason
#| echo: true

plotMergedBySeason <- function(d, r, minYear = 2000, maxYear = 2017) {
  ggplot(d |> filter(riverGG == r, year %in% c(minYear:maxYear), !is.na(seasonGG)), 
         aes(temperature, log10(flowByRiverm3s), 
             color = factor(year), group = year
             )
         ) +
    #geom_point() +
    geom_smooth(se = FALSE) +
    labs(x = "Temperature (C)", y = "log10 of Flow") +
    scale_color_discrete() +
    ggtitle(r) +
    facet_wrap(~seasonGG, scales = 'free')
}

```

::: panel-tabset
#### West Brook

```{r}
#| label: plotMergedBySeason_wb

plotMergedBySeason(envIn_propNeg, "West Brook", 2000, 2017)
```

#### Open Large

```{r}
#| label: plotMergedBySeason_ol

plotMergedBySeason(envIn_propNeg, "Open Large", 2002, 2017)
```

#### Open Small

```{r}
#| label: plotMergedBySeason_os

plotMergedBySeason(envIn_propNeg, "Open Small", 2002, 2017)
```

#### Isolated small

```{r}
#| label: plotMergedBySeason_is

plotMergedBySeason(envIn_propNeg, "Isolated Small", 2002, 2017)
```
:::

TO ADD: low flow high density important? Look at interaction btw
flow:density

### F/T cycles with labels

Plot F/T cycles by river, season, species with label for propNeg and
conspecific and all-species abundances

```{r}
#| label: plotMergedBySeasonByEverything

plotMergedBySeasonByEverything <- function(d, d2, r, s, spp, minYear = 2000, maxYear = 2017){
  ggplot(d |> 
           filter(riverGG == r,
                  year %in% c(minYear:maxYear), 
                  !is.na(seasonGG), 
                  seasonGG == s, 
                  speciesGG == spp), 
         aes(temperature, log10(flowByRiverm3s))
         ) +
    geom_smooth(se = FALSE, color = "darkgrey") +
    geom_point(aes(color = yday)) +
    geom_text(aes(
      x = minTemp + 3.0, 
      y = log10(minFlow) - 1, 
      label = paste("propNeg=",round(propNeg, 2)) 
      #size = propNeg
      ), 
        data = d2|> 
          filter(riverGG == r, 
                 year %in% c(minYear:maxYear),
                 seasonGG == s, 
                 speciesGG == spp)
        ) +
    geom_text(aes(
      x = minTemp + 3.0,
      y = log10(minFlow) - 0.3,
      label = paste(paste("nIndAllSpp=", nInd), paste("nIndBySpp=", nIndBySpp), sep = "\n")
      ),
        data = d2|>
          filter(riverGG == r,
                 year %in% c(minYear:maxYear),
                 seasonGG == s,
                 speciesGG == spp)
        ) +
    labs(x = "Temperature (c)", y = "log10 Flow") +
    ggtitle(paste(r, s, spp, sep = ", ")) +
  facet_wrap(~year)
}

```

#### West Brook

::: panel-tabset
#### Spring

```{r}
#| label: WBPanelSpring
#| fig-width: 13
#| fig-height: 10
#| 
plotMergedBySeasonByEverything(envIn_propNeg, propNegLabels, "West Brook", "Spring", "Brook trout", 2000, 2017)
plotMergedBySeasonByEverything(envIn_propNeg, propNegLabels, "West Brook", "Spring", "Brown trout", 2000, 2017)
plotMergedBySeasonByEverything(envIn_propNeg, propNegLabels, "West Brook", "Spring", "Atlantic salmon", 2000, 2017)


```

#### Summer

```{r}
#| label: WBPanelSummer
#| fig-width: 13
#| fig-height: 10
plotMergedBySeasonByEverything(envIn_propNeg, propNegLabels, "West Brook", "Summer", "Brook trout", 2000, 2017)
plotMergedBySeasonByEverything(envIn_propNeg, propNegLabels, "West Brook", "Summer", "Brown trout", 2000, 2017)
plotMergedBySeasonByEverything(envIn_propNeg, propNegLabels, "West Brook", "Summer", "Atlantic salmon", 2000, 2017)
```

#### Autumn

```{r}
#| label: WBPanelAutumn
#| fig-width: 13
#| fig-height: 10
#| 
plotMergedBySeasonByEverything(envIn_propNeg, propNegLabels, "West Brook", "Autumn", "Brook trout", 2000, 2017)
plotMergedBySeasonByEverything(envIn_propNeg, propNegLabels, "West Brook", "Autumn", "Brown trout", 2000, 2017)
plotMergedBySeasonByEverything(envIn_propNeg, propNegLabels, "West Brook", "Autumn", "Atlantic salmon", 2000, 2017)


```

#### Winter

```{r}
#| label: WBPanelWinter
#| fig-width: 13
#| fig-height: 10
plotMergedBySeasonByEverything(envIn_propNeg, propNegLabels, "West Brook", "Winter", "Brook trout", 2000, 2017)
plotMergedBySeasonByEverything(envIn_propNeg, propNegLabels, "West Brook", "Winter", "Brown trout", 2000, 2017)
plotMergedBySeasonByEverything(envIn_propNeg, propNegLabels, "West Brook", "Winter", "Atlantic salmon", 2000, 2017)
```
:::

#### Open Large

::: panel-tabset
#### Spring

```{r}
#| label: OLpanelSpring
#| fig-width: 13
#| fig-height: 10
#| 
plotMergedBySeasonByEverything(envIn_propNeg, propNegLabels, "Open Large", "Spring", "Brook trout", 2000, 2017)
plotMergedBySeasonByEverything(envIn_propNeg, propNegLabels, "Open Large", "Spring", "Brown trout", 2000, 2017)
plotMergedBySeasonByEverything(envIn_propNeg, propNegLabels, "Open Large", "Spring", "Atlantic salmon", 2000, 2017)


```

#### Summer

```{r}
#| label: OLPanelSummer
#| fig-width: 13
#| fig-height: 10
plotMergedBySeasonByEverything(envIn_propNeg, propNegLabels, "Open Large", "Summer", "Brook trout", 2000, 2017)
plotMergedBySeasonByEverything(envIn_propNeg, propNegLabels, "Open Large", "Summer", "Brown trout", 2000, 2017)
plotMergedBySeasonByEverything(envIn_propNeg, propNegLabels, "Open Large", "Summer", "Atlantic salmon", 2000, 2017)
```

#### Autumn

```{r}
#| label: OLPanelAutumn
#| fig-width: 13
#| fig-height: 10
#| 
plotMergedBySeasonByEverything(envIn_propNeg, propNegLabels, "Open Large", "Autumn", "Brook trout", 2000, 2017)
plotMergedBySeasonByEverything(envIn_propNeg, propNegLabels, "Open Large", "Autumn", "Brown trout", 2000, 2017)
plotMergedBySeasonByEverything(envIn_propNeg, propNegLabels, "Open Large", "Autumn", "Atlantic salmon", 2000, 2017)


```

#### Winter

```{r}
#| label: OLPanelWinter
#| fig-width: 13
#| fig-height: 10
plotMergedBySeasonByEverything(envIn_propNeg, propNegLabels, "Open Large", "Winter", "Brook trout", 2000, 2017)
plotMergedBySeasonByEverything(envIn_propNeg, propNegLabels, "Open Large", "Winter", "Brown trout", 2000, 2017)
plotMergedBySeasonByEverything(envIn_propNeg, propNegLabels, "Open Large", "Winter", "Atlantic salmon", 2000, 2017)
```
:::

#### Open Small

::: panel-tabset
#### Spring

```{r}
#| label: OSPanelSpring
#| fig-width: 13
#| fig-height: 10
#| 
plotMergedBySeasonByEverything(envIn_propNeg, propNegLabels, "Open Small", "Spring", "Brook trout", 2000, 2017)
plotMergedBySeasonByEverything(envIn_propNeg, propNegLabels, "Open Small", "Spring", "Brown trout", 2000, 2017)
```

#### Summer

```{r}
#| label: OSPanelSummer
#| fig-width: 13
#| fig-height: 10
plotMergedBySeasonByEverything(envIn_propNeg, propNegLabels, "Open Small", "Summer", "Brook trout", 2000, 2017)
plotMergedBySeasonByEverything(envIn_propNeg, propNegLabels, "Open Small", "Summer", "Brown trout", 2000, 2017)

```

#### Autumn

```{r}
#| label: OSPanelAutumn
#| fig-width: 13
#| fig-height: 10
#| 
plotMergedBySeasonByEverything(envIn_propNeg, propNegLabels, "Open Small", "Autumn", "Brook trout", 2000, 2017)
plotMergedBySeasonByEverything(envIn_propNeg, propNegLabels, "Open Small", "Autumn", "Brown trout", 2000, 2017)

```

#### Winter

```{r}
#| label: OSPanelWinter
#| fig-width: 13
#| fig-height: 10
plotMergedBySeasonByEverything(envIn_propNeg, propNegLabels, "Open Small", "Winter", "Brook trout", 2000, 2017)
plotMergedBySeasonByEverything(envIn_propNeg, propNegLabels, "Open Small", "Winter", "Brown trout", 2000, 2017)

```
:::

#### Isolated Small

::: panel-tabset
#### Spring

```{r}
#| label: ISPanelSpring
#| fig-width: 13
#| fig-height: 10
#| 
plotMergedBySeasonByEverything(envIn_propNeg, propNegLabels, "Isolated Small", "Spring", "Brook trout", 2000, 2017)
```

#### Summer

```{r}
#| label: ISPanelSummer
#| fig-width: 13
#| fig-height: 10
plotMergedBySeasonByEverything(envIn_propNeg, propNegLabels, "Isolated Small", "Summer", "Brook trout", 2000, 2017)

```

#### Autumn

```{r}
#| label: ISPanelAutumn
#| fig-width: 13
#| fig-height: 10
#| 
plotMergedBySeasonByEverything(envIn_propNeg, propNegLabels, "Isolated Small", "Autumn", "Brook trout", 2000, 2017)

```

#### Winter

```{r}
#| label: ISPanelWinter
#| fig-width: 13
#| fig-height: 10
plotMergedBySeasonByEverything(envIn_propNeg, propNegLabels, "Isolated Small", "Winter", "Brook trout", 2000, 2017)

```
:::

## Consequences of negative growth

### Growth and last observation

Relationship between individual growth and length of subsequent
observations (how many more observations before the last observatoin)
for individual fish.

```{r}
#| label: negGrowth_samplesLeft
plotGrowthSamplesLeft <- function(d, spp, r){
  ggplot(d |> filter(speciesGG == spp, riverGG == r, year < 2011, samplesBeforeLast >= 0), aes(grWeight, samplesBeforeLast)) +
    geom_point(alpha = 0.3) +
    geom_smooth(method = "lm", se = FALSE) +
    geom_vline(xintercept = 0) +
    ggtitle(paste(spp, r, sep = ", ")) +
    facet_grid(seasonGG~year)
}

plotGrowthSamplesLeft(cd1, "Brook trout", "West Brook")
plotGrowthSamplesLeft(cd1, "Brook trout", "Open Large")
plotGrowthSamplesLeft(cd1, "Brook trout", "Open Small")
plotGrowthSamplesLeft(cd1, "Brook trout", "Isolated Small")

plotGrowthSamplesLeft(cd1, "Brown trout", "West Brook")
```

```{r}
#| label: negGrowth_samplesLeftNoYear
plotGrowthSamplesLeftNoYear <- function(d, spp, r){
  ggplot(d |> filter(speciesGG == spp, riverGG == r, year < 2011, samplesBeforeLast >= 0), aes(grWeight, samplesBeforeLast)) +
    geom_point(alpha = 0.3) +
    geom_smooth(method = "lm", se = FALSE) +
    geom_vline(xintercept = 0) +
    ggtitle(paste(spp, r, sep = ", ")) +
    labs(x = "Individual growth in mass", y = "Number of samples before last observation") +
    facet_wrap(~seasonGG)
}

plotGrowthSamplesLeftNoYear(cd1, "Brook trout", "West Brook")
plotGrowthSamplesLeftNoYear(cd1, "Brook trout", "Open Large")
plotGrowthSamplesLeftNoYear(cd1, "Brook trout", "Open Small")
plotGrowthSamplesLeftNoYear(cd1, "Brook trout", "Isolated Small")

plotGrowthSamplesLeftNoYear(cd1, "Brown trout", "West Brook")
plotGrowthSamplesLeftNoYear(cd1, "Atlantic salmon", "West Brook")
```

```{r}
#| label: negGrowth_means

plotGR_sampsBL <- function(d, spp, r) {
  ggplot(d |> filter(n > 5, speciesGG %in% spp, riverGG %in% r), aes(meanGR, samplesBeforeLast, color = negGrowth)) + 
    geom_point() +
    geom_smooth(method = "lm", se = FALSE) +
    ggtitle(paste(spp, r, sep = ", ")) +
    labs(x = "Individual growth in mass", y = "Number of samples before last observation") +
    facet_grid(speciesGG~seasonGG+riverGG, scales = "free")
}

plotGR_sampsBL(negGr_beforeLast, c("Brook trout","Brown trout", "Atlantic salmon"), c("West Brook", "Open Large"))


plotGR_sampsBL2 <- function(d, spp, r) {
  ggplot(d |> filter(n > 5, speciesGG %in% spp, riverGG %in% r), aes(meanGR, samplesBeforeLast, color = speciesGG)) + 
    geom_point() +
    geom_smooth(method = "lm", se = FALSE) +
    scale_color_discrete() +
    ggtitle(paste(spp, r, sep = ", ")) +
    labs(x = "Individual growth in mass", y = "Number of samples before last observation") +
    facet_grid(riverGG~negGrowth+seasonGG, scales = "free")
}

plotGR_sampsBL2(negGr_beforeLast, c("Brook trout","Brown trout", "Atlantic salmon"), c("West Brook", "Open Large"))

plotGR_sampsBL3 <- function(d, spp, r, ng) {
  ggplot(d |> filter(n > 5, speciesGG %in% spp, riverGG %in% r, negGrowth %in% ng), aes(meanGR, samplesBeforeLast, color = speciesGG)) + 
    geom_point() +
    geom_smooth(method = "lm", se = FALSE) +
    scale_color_discrete() +
    ggtitle(paste(spp, r, ng, sep = ", ")) +
    labs(x = "Individual growth in mass", y = "Number of samples before last observation") +
    facet_wrap(negGrowth~seasonGG, scales = "free")
}

plotGR_sampsBL3(negGr_beforeLast, c("Brook trout","Brown trout", "Atlantic salmon"), c("West Brook"), FALSE)

plotGR_sampsBL3(negGr_beforeLast, c("Brook trout","Brown trout", "Atlantic salmon"), c("West Brook"), TRUE)
```

TO ADD: mean samplesBeforeLast by negGR TO ADD: knownZ=0 vs knownZ==1,
growth in previous interval

```{r}

# cd1_target |> 
#       group_by(speciesGG, riverGG, seasonGG, samplesBeforeLast, negGrowth) |> 

```

### Growth and movement distance

TO add: break avg move into negGR/not negGR

```{r}
#| label: growthMovement

plotGR_moveDist <- function(d, spp, r) {
  ggplot(d |> filter(speciesGG %in% spp, riverGG %in% r), aes(grWeight, distMoved, color = speciesGG)) + 
    geom_point(alpha = 0.2) +
    geom_smooth(method = "lm", se = FALSE, color = 'darkgrey') +
    scale_color_discrete() +
    ggtitle(paste(spp, r, sep = ", ")) +
    labs(x = "Individual growth in mass", y = "Movement distance") +
    facet_grid(speciesGG~seasonGG, scales = "free")
}

plotGR_moveDist(cd1, c("Brook trout","Brown trout","Atlantic salmon"), c("West Brook"))


```

```{r}
#| label: growthMovement2

plotGR_moveDist2 <- function(d, spp, r) {
  ggplot(d |> filter(speciesGG %in% spp, riverGG %in% r), aes(grWeight, distMoved, color = speciesGG)) + 
    geom_point(alpha = 0.2) +
    geom_smooth(method = "lm", se = FALSE, color = 'darkgrey') +
    scale_color_discrete() +
    ggtitle(paste(spp, r, sep = ", ")) +
    labs(x = "Individual growth in mass", y = "Movement distance") +
    facet_grid(riverGG~seasonGG, scales = "free")
}

plotGR_moveDist2(cd1, c("Brook trout"), c("West Brook", "Open Large", "Open Small", "Isolated Small"))

plotGR_moveDist2(cd1, c("Brown trout"), c("West Brook", "Open Large", "Open Small"))
```

## Machine Learning models

Using xgBoost to explore patterns in:\
- Binary, lose mass:gain mass\
- Regression, growth in mass

### Binary model with negative growth, yes:no

Not yet in targets....

```{r}
#| label: rf
# http://uc-r.github.io/gbm_regression


dML <- cd1 |> 
  filter(speciesGG == "Brook trout", !is.na(grWeight)) |> 
  mutate(age = year - cohort) |> 
  dplyr::select(riverGG, seasonGG, observedWeight, cohort, age, meanTemperature, meanFlowByRiver, negGrowth) 

# variable names
features <- setdiff(names(dML), "negGrowth")

# Create the treatment plan from the training data
treatplan <- vtreat::designTreatmentsZ(dML, features, verbose = FALSE)

# Get the "clean" variable names from the scoreFrame
new_vars <- treatplan %>%
  magrittr::use_series(scoreFrame) %>%        
  dplyr::filter(code %in% c("clean", "lev")) %>% 
  magrittr::use_series(varName)     

# Prepare the training data
features_train <- vtreat::prepare(treatplan, dML, varRestriction = new_vars) %>% as.matrix()
response_train <- dML$negGrowth

# Prepare the test data
features_test <- vtreat::prepare(treatplan, dML, varRestriction = new_vars) %>% as.matrix()
response_test <- dML$negGrowth

# dimensions of one-hot encoded data
dim(features_train)
dim(features_test)
```

#### Run first model

```{r}
#| label: runML1
#| 
# reproducibility
set.seed(123)

xgb.fit1 <- xgb.cv(
  data = features_train,
  label = response_train,
  nrounds = 10000,
  nfold = 5,
  objective = "binary:logistic", #"reg:linear",  # for regression models
  verbose = 0,               # silent,
  early_stopping_rounds = 15 # stop if no improvement for 10 consecutive trees
)

# get number of trees that minimize error
xgb.fit1$evaluation_log %>%
  dplyr::summarise(
    ntrees.train = which(train_logloss_mean == min(train_logloss_mean))[1],
    logloss.train   = min(train_logloss_mean),
    ntrees.test  = which(test_logloss_mean == min(test_logloss_mean))[1],
    logloss.test   = min(test_logloss_mean),
  )

# plot error vs number trees
ggplot(xgb.fit1$evaluation_log) +
  geom_line(aes(iter, train_logloss_mean), color = "red") +
  geom_line(aes(iter, test_logloss_mean), color = "blue")
```

#### Add early stopping

```{r}
#| label: runML2

## tune the model
# create parameter list
params <- list(
  eta = .1,
  max_depth = 5,
  min_child_weight = 2,
  subsample = .8,
  colsample_bytree = .9
)

xgb.fit2 <- xgb.cv(
  params = params,
  data = features_train,
  label = response_train,
  nrounds = 20000,
  nfold = 5,
  objective = "binary:logistic", #"reg:linear",  # for regression models
  verbose = 0,               # silent,
  early_stopping_rounds = 10 # stop if no improvement for 10 consecutive trees
)

xgb.fit2$evaluation_log %>%
  dplyr::summarise(
    ntrees.train = which(train_logloss_mean == min(train_logloss_mean))[1],
    logloss.train   = min(train_logloss_mean),
    ntrees.test  = which(test_logloss_mean == min(test_logloss_mean))[1],
    logloss.test   = min(test_logloss_mean),
  )

# plot error vs number trees
ggplot(xgb.fit2$evaluation_log) +
  geom_line(aes(iter, train_logloss_mean), color = "red") +
  geom_line(aes(iter, test_logloss_mean), color = "blue")

```

#### Hyperparameter search grid

```{r}
#| label: runML_HP


# create hyperparameter grid
hyper_grid <- expand.grid(
  eta = c(.01, .05, .1, .3),
  max_depth = c(1, 3, 5, 7),
  min_child_weight = c(1, 3, 5, 7),
  subsample = c(.65, .8, 1), 
  colsample_bytree = c(.8, .9, 1),
  optimal_trees = 0,               # a place to dump results
  min_RMSE = 0                     # a place to dump results
)
nrow(hyper_grid)

  
# reproducibility
  set.seed(123)
#   
# # grid search 
# for(i in 1:nrow(hyper_grid)) {
#   print(i)
#   # create parameter list
#   params <- list(
#     eta = hyper_grid$eta[i],
#     max_depth = hyper_grid$max_depth[i],
#     min_child_weight = hyper_grid$min_child_weight[i],
#     subsample = hyper_grid$subsample[i],
#     colsample_bytree = hyper_grid$colsample_bytree[i]
#   )
# 
#   
#   # train model
#   xgb.tune <- xgb.cv(
#     params = params,
#     data = features_train,
#     label = response_train,
#     nrounds = 5000,
#     nfold = 5,
#     objective = "binary:logistic", #"reg:linear",  # for regression models
#     verbose = 0,               # silent,
#     early_stopping_rounds = 20 # stop if no improvement for 10 consecutive trees
#   )
#   
#   # add min training error and trees to grid
#   hyper_grid$optimal_trees[i] <- which.min(xgb.tune$evaluation_log$test_logloss_mean)
#   hyper_grid$min_RMSE[i] <- min(xgb.tune$evaluation_log$test_logloss_mean)
# }
# 
# hyper_grid %>%
#   dplyr::arrange(min_RMSE) %>%
#   head(15)
  
#       eta max_depth min_child_weight subsample colsample_bytree optimal_trees  min_RMSE
# 1  0.01         5                1      0.80              0.8          1082 0.2464962
# 2  0.01         5                1      0.65              0.8          1042 0.2466260
# 3  0.01         5                1      0.80              0.9          1115 0.2467519
# 4  0.01         7                1      0.80              0.8           676 0.2467640
# 5  0.01         7                1      0.65              0.8           671 0.2468416
# 6  0.05         5                1      0.65              0.9           217 0.2470321
# 7  0.01         7                3      0.80              0.8           676 0.2470520
# 8  0.01         5                3      0.80              0.8          1115 0.2470706
# 9  0.01         5                3      0.80              0.9          1089 0.2470969
# 10 0.10         5                3      1.00              0.9           158 0.2471294
# 11 0.05         7                3      0.80              1.0           139 0.2471420
# 12 0.01         5                1      0.65              0.9           960 0.2471568
# 13 0.01         5                1      0.80              1.0          1023 0.2471608
# 14 0.01         7                3      0.65              0.8           671 0.2472037
# 15 0.10         5                1      0.80              0.8            95 0.2472090
```

#### Fit the top model

```{r}
#| label: fitBinary

# parameter list - from the top model of the grid search above
params <- list(
  eta = 0.01,
  max_depth = 5,
  min_child_weight = 1,
  subsample = 0.8,
  colsample_bytree = 0.8
)

# train final model
xgb.fit.final <- xgboost(
  params = params,
  data = features_train,
  label = response_train,
  nrounds = 1082,
  objective = "binary:logistic",
  verbose = 0,
  early_stopping_rounds = 25 # stop if no improvement for 10 consecutive trees
)

```

#### Variable importance plot

```{r}
#| label: vipBinary

# create importance matrix
importance_matrix <- xgb.importance(model = xgb.fit.final)

# variable importance plot
xgb.plot.importance(importance_matrix, top_n = 10, measure = "Gain")

```

#### Partial dependence plots

```{r}
#| label: pdpBinary

pW <- xgb.fit.final %>%
  pdp::partial(pred.var = "observedWeight", n.trees = 1082, grid.resolution = 100, train = features_train) %>%
  autoplot(rug = TRUE, train = features_train) 
  #scale_y_discrete() +
  #ggtitle("PDP")

pFlow <- xgb.fit.final %>%
  pdp::partial(pred.var = "meanFlowByRiver", n.trees = 1082, grid.resolution = 100, train = features_train) %>%
  autoplot(rug = TRUE, train = features_train)

pTemp <- xgb.fit.final %>%
  pdp::partial(pred.var = "meanTemperature", n.trees = 1082, grid.resolution = 100, train = features_train) %>%
  autoplot(rug = TRUE, train = features_train)

pAge <- xgb.fit.final %>%
  pdp::partial(pred.var = "age", n.trees = 1082, grid.resolution = 100, train = features_train) %>%
  autoplot(rug = TRUE, train = features_train)

pCohort <- xgb.fit.final %>%
  pdp::partial(pred.var = "cohort", n.trees = 1082, grid.resolution = 100, train = features_train) %>%
  autoplot(rug = TRUE, train = features_train)


gridExtra::grid.arrange(pW, pFlow, pTemp, pAge, pCohort, nrow = 3)
```

#### Lime

```{r}
#| label: limeBinary

# get a few observations to perform local interpretation on
local_obs <- dML[1:3, ]

# one-hot encode the local observations to be assessed.
local_obs_onehot <- vtreat::prepare(treatplan, local_obs, varRestriction = new_vars)

# apply LIME
explainer <- lime(data.frame(features_train), xgb.fit.final, quantile_bins = FALSE)
explanation <- lime::explain(local_obs_onehot, explainer, n_features = 5, n_labels = 2)
plot_features(explanation)
```

### Growth model, continuous growth

can we add season, river interactions? or do separate analyses?

```{r}
#| label: readInTargets_W
#| cache: false
library(targets)
dML_W <- tar_read(dML_W_target)
treatplan_W <- tar_read(treatplan_W_target)
features_train_W <- tar_read(features_train_W_target)
new_vars_W <- tar_read(new_vars_W_target)

# dimensions of one-hot encoded data
features_train_W <- tar_read(features_train_W_target)
features_test_W <- tar_read(features_test_W_target)

response_train_W <- tar_read(response_train_W_target)
response_test_W <- tar_read(response_test_W_target)

# run from targets
hyper_grid_W <- tar_read(hyper_grid_W_target)
topModel_W <- tar_read(topModel_W_target)


finalModels_W <- tar_read(finalModels_W_target)

modelToUse <- 1
```

#### Run first model

```{r}
#| label: runML1_W
#| 
# reproducibility
set.seed(1234)

xgb.fit1_W <- xgb.cv(
  data = features_train_W,
  label = response_train_W,
  nrounds = 500,
  nfold = 5,
  objective = "reg:squarederror",  # for regression models
  verbose = 0,               # silent,
  early_stopping_rounds = 10
)

# get number of trees that minimize error
xgb.fit1_W$evaluation_log %>%
  dplyr::summarise(
    ntrees.train = which(train_rmse_mean == min(train_rmse_mean))[1],
    rmse.train   = min(train_rmse_mean),
    ntrees.test  = which(test_rmse_mean == min(test_rmse_mean))[1],
    rmse.test   = min(test_rmse_mean),
  )

# plot error vs number trees
ggplot(xgb.fit1_W$evaluation_log) +
  geom_line(aes(iter, train_rmse_mean), color = "red") +
  geom_line(aes(iter, test_rmse_mean), color = "blue")
```

#### Hyperparameter search grid

```{r}
#| label: runML_HP_W


# create hyperparameter grid
# original grid - using a smaller one in targets
# hyper_grid_W0 <- expand.grid(
#   eta = c(.01, .05, .1, .3),
#   max_depth = c(1, 3, 5, 7),
#   min_child_weight = c(1, 3, 5, 7),
#   subsample = c(.65, .8, 1), 
#   colsample_bytree = c(.8, .9, 1),
#   optimal_trees = 0,               # a place to dump results
#   min_RMSE = 0                     # a place to dump results
# )
# nrow(hyper_grid_W0)



# Original run using hyper_grid_W0 
# eta max_depth min_child_weight
# subsample colsample_bytree optimal_trees min_RMSE 1 0.01 7 5 0.65 0.8
# 1376 0.001847600 2 0.05 7 3 0.80 0.9 292 0.001848465 3 0.01 7 7 0.80 0.9
# 1382 0.001849086 4 0.01 5 1 0.80 0.8 2399 0.001849250 5 0.05 7 3 0.80
# 0.8 281 0.001849258 6 0.01 5 3 0.65 0.8 2147 0.001849905 7 0.05 5 1 0.65
# 0.8 459 0.001849931 8 0.01 5 5 0.80 0.9 2296 0.001850049 9 0.01 7 5 0.65
# 0.9 1402 0.001850246 10 0.01 7 7 0.65 0.8 1382 0.001850544 11 0.01 7 3
# 0.65 0.8 1384 0.001850704 12 0.05 3 3 0.80 0.8 1509 0.001850763 13 0.01
# 7 7 0.80 0.8 1453 0.001850975 14 0.01 7 1 0.80 1.0 1281 0.001851069 15
# 0.01 7 1 0.80 0.8 1473 0.001851078


```


#### Variable importance plot

```{r}
#| label: vipContinuous_W
#| cache: false

#importance_matrix_W <- tar_read(importance_matrix_W_target)

# variable importance plot


```

#### Partial dependence plots

```{r}
#| label: pdpContinuous_W

create_xgb_plots <- function(d = finalModels_W, f = features_train_W, modelToUse = modelToUse) {
  
  imp <- xgb.ggplot.importance(d$importanceMatrix[[modelToUse]], top_n = 10, measure = "Gain") + 
    ggtitle(paste0("Model #", modelToUse))

  pW_W <- d$xgb[[modelToUse]] %>%
    pdp::partial(pred.var = "observedWeight", n.trees = 1082, grid.resolution = 100, train = f) %>%
    autoplot(rug = TRUE, train = f) 
  
  pFlow_W <- d$xgb[[modelToUse]] %>%
    pdp::partial(pred.var = "meanFlowByRiver", n.trees = 1082, grid.resolution = 100, train = f) %>%
    autoplot(rug = TRUE, train = f)
  
  pTemp_W <- d$xgb[[modelToUse]] %>%
    pdp::partial(pred.var = "meanTemperature", n.trees = 1082, grid.resolution = 100, train = f) %>%
    autoplot(rug = TRUE, train = f)
  
  pAge_W <- d$xgb[[modelToUse]] %>%
    pdp::partial(pred.var = "age", n.trees = 1082, grid.resolution = 100, train = f) %>%
    autoplot(rug = TRUE, train = f)
  
  pCohort_W <- d$xgb[[modelToUse]] %>%
    pdp::partial(pred.var = "cohort", n.trees = 1082, grid.resolution = 100, train = f) %>%
    autoplot(rug = TRUE, train = f)


  partials <- gridExtra::grid.arrange(pW_W, pFlow_W, pTemp_W, pAge_W, pCohort_W, nrow = 3, top=textGrob(paste0("Model #", modelToUse)))
  
 return(list(imp, partials)) 
}

create_xgb_plots(finalModels_W, features_train_W, modelToUse = 1)
create_xgb_plots(finalModels_W, features_train_W, modelToUse = 2)
create_xgb_plots(finalModels_W, features_train_W, modelToUse = 3)
create_xgb_plots(finalModels_W, features_train_W, modelToUse = 4)
create_xgb_plots(finalModels_W, features_train_W, modelToUse = 5)
create_xgb_plots(finalModels_W, features_train_W, modelToUse = 15)
```

#### Lime

```{r}
#| label: limeContinuous_W

# get a few observations to perform local interpretation on
local_obs_W <- dML_W[1:3, ]

# one-hot encode the local observations to be assessed.
local_obs_onehot_W <- vtreat::prepare(treatplan_W, local_obs_W, varRestriction = new_vars_W)

# apply LIME
explainer_W <- lime(data.frame(features_train_W), finalModels_W$xgb[[modelToUse]], quantile_bins = FALSE)
explanation_W <- lime::explain(local_obs_onehot_W, explainer_W, n_features = 5, n_labels = 2)
plot_features(explanation_W)
```

#### Predict

```{r}
#| label: predict_W


createPredictions <- function(d = finalModels_W, f = features_test_W, r = response_test_W, modelToUse = modelToUse) {
  
  pred_W <- predict(d$xgb[[modelToUse]], f)
  rmse <- caret::RMSE(pred_W, r)
  r2 <- caret::R2(pred_W, r)
  
  obsPred_W <- bind_cols(obs = r, pred = pred_W, f)
  
  return(list(r2 = r2, rmse = rmse, obsPred_W = obsPred_W))
}

```

```{r}
#| label: obsPred_W
#| 

predsOP_W <- createPredictions(finalModels_W, features_test_W, response_test_W, modelToUse) 

ggplot(predsOP_W$obsPred_W, aes(obs, pred)) +
  geom_point(alpha = 0.2) +
  labs(x = "Observed growth rate", y = "Predicted growth rate") +
  geom_smooth(method = "lm", se = FALSE) +
  ggtitle(paste0("r2 = ", round(predsOP_W$r2, 2), " RMSE = ", round(predsOP_W$rmse, 5)))

ggplot(predsOP_W$obsPred_W, aes(obs, pred)) +
  geom_point(alpha = 0.2) +
  labs(x = "Observed growth rate", y = "Predicted growth rate") +
  geom_smooth(method = "lm", se = FALSE) +
  facet_wrap(~cohort)
```

```{r}
#| label: UsingGridInputs
#| 
renameLev <- function(d) {
  d |>  
    str_split("_lev_x_") |> 
    map(~.x[2]) |> 
    unlist()
}

createPredictionsGrid <- function(d = finalModels_W, pg = gridPred, modelToUse = modelToUse){
  #, f = features_test_W, r = response_test_W, nSeq = 5, modelToUse = modelToUse) {
  
  predGrid <- predict(d$xgb[[modelToUse]], pg)
  gridPredOut <- bind_cols(pg, pred = predGrid) |> 
    rename_with(
      renameLev,  
      contains("_lev_x_")
    ) |>  
    rowwise() |> 
    mutate(
      sumRiver = sum(Isolated_Small + Open_Large + Open_Small + West_Brook),
      sumSeason = sum(Autumn + Spring + Summer + Winter)
    ) |>   
    filter(
      sumRiver == 1,
      sumSeason == 1
    ) |>   
    mutate(
      across(c('Isolated_Small':'Winter'), ~.x == 1)
    )

  return(gridPredOut)
}



nSeq <- 5
xGrid <- expand_grid(
    observedWeight = 40, #seq(min(f[,"observedWeight"]), max(f[,"observedWeight"]), length.out = nSeq),
    cohort = 2010, #seq(min(f[,"cohort"]), max(f[,"cohort"]), length.out = nSeq),
    age = 1, #seq(min(f[,"age"]), max(f[,"age"]), length.out = nSeq),
    meanTemperature = seq(min(features_test_W[,"meanTemperature"]), max(features_test_W[,"meanTemperature"]), length.out = nSeq * 3),
    meanFlowByRiver = seq(min(features_test_W[,"meanFlowByRiver"]), max(features_test_W[,"meanFlowByRiver"]), length.out = nSeq * 3),
    riverGG_lev_x_Isolated_Small = c(0, 1),
    riverGG_lev_x_Open_Large = c(0, 1),
    riverGG_lev_x_Open_Small = c(0, 1),
    riverGG_lev_x_West_Brook = c(0, 1),
    seasonGG_lev_x_Autumn = c(0, 1),
    seasonGG_lev_x_Spring = c(0, 1),
    seasonGG_lev_x_Summer = c(0, 1),
    seasonGG_lev_x_Winter = c(0, 1)
  ) |> 
  as.matrix()


preds_W <- createPredictionsGrid(
  finalModels_W, 
  xGrid,
  modelToUse
) 


ggplot(preds_W |> 
         filter(
           #Isolated_Small, Open_Large, Open_Small, West_Brook, 
           #Autumn, Spring, Summer, Winter
           ), 
       aes(meanFlowByRiver, pred, color = meanTemperature, group = meanTemperature)) +
  geom_point(alpha = 0.2) +
  labs(x = "Stream flow", y = "Predicted growth rate") +
  geom_smooth(se = FALSE) +
  geom_hline(yintercept = 0) +
  facet_grid(Isolated_Small+Open_Large+Open_Small+West_Brook ~ Spring+Summer+Autumn+Winter)

# use seasonal zScores for environmetnal data
```


#### When and where do fish have predicted negative growth?
```{r}
#| label: negPredGrowth_W

predNeg <- preds_W |> filter(pred < 0.000)
predNeg |> 
  count(Isolated_Small, Open_Large, Open_Small, West_Brook, Autumn, Spring, Summer, Winter) |> 
  arrange(desc(n))

predNeg |> 
  count(Autumn, Spring, Summer, Winter) |> 
  arrange(desc(n))

predNeg |> 
  count(Isolated_Small, Open_Large, Open_Small, West_Brook) |> 
  arrange(desc(n))

```

## GAM for growth
```{r}
#| label: gamGrowthRaw

#riverGG, seasonGG, observedWeight, cohort, age, meanTemperature, meanFlowByRiver, negGrowth


kValue <- 4 

ggplot(dGAM |> filter(speciesGG != "Atlantic salmon"), aes(flowByRiverS, grWeight, group = speciesGG, color = speciesGG)) +
  geom_point(alpha = 0.2) +
  geom_smooth(method = "gam", 
              formula = y ~ s(x, bs = "cs", fx = TRUE, k = kValue), 
              se = FALSE) +
  facet_grid(seasonGG~riverGG)

ggplot(dGAM |> filter(speciesGG != "Atlantic salmon"), aes(flowByRiverSR, grWeight, group = speciesGG, color = speciesGG)) +
  geom_point(alpha = 0.1) +
  geom_hline(yintercept = 0) +
  geom_smooth(method = "gam", 
              formula = y ~ s(x, bs = "cs", fx = TRUE, k = kValue), 
              se = FALSE) +
  scale_color_discrete() +
  facet_grid(seasonGG~riverGG, scales = "free")


ggplot(dGAM |> filter(speciesGG != "Atlantic salmon"), aes(tempSR, grWeight, group = speciesGG, color = speciesGG)) +
  geom_point(alpha = 0.1) +
  geom_hline(yintercept = 0) +
  geom_smooth(method = "gam", 
              formula = y ~ s(x, bs = "cs", fx = TRUE, k = kValue), 
              se = FALSE) +
  scale_color_discrete() +
  facet_grid(seasonGG~riverGG, scales = "free")
```

### Linear models for growth  
Initial test of model structure for GAMs
```{r}
#| label: linearModelGrowth

runLinearModels <- function(d, speciesIn) {
  models =
    tibble(model = 
      c(
          'grWeight ~ (riverGG * seasonGG * tempSR * flowByRiverSR * observedWeight + cohortF * ageF)',
          'grWeight ~ (riverGG * seasonGG * tempSR * flowByRiverSR * observedWeight + cohortF)',
          'grWeight ~ (riverGG * seasonGG * tempSR * flowByRiverSR + cohortF * ageF)',
          'grWeight ~ (riverGG * seasonGG * tempSR * flowByRiverSR + cohortF)',
          'grWeight ~ (riverGG * seasonGG + tempSR * flowByRiverSR + cohortF * ageF)',
          'grWeight ~ (riverGG * seasonGG + tempSR + flowByRiverSR + cohortF * ageF)'
       )
    ) |>
    mutate(
      fit = map(model, ~lm(.x, data = d |> filter(speciesGG == speciesIn))),
      glanced = map(fit, broom::glance),
      tidied = map(fit, broom::tidy)
    )
  return(models)
}


```
#### Brook trout  
Most complex model has lowest AIC
```{r}
#| label: linBKT
#| 
speciesIn <- "Brook trout"
modLinearBKT <- runLinearModels(dGAM, speciesIn)

modLinearBKT |> 
    unnest(glanced) |>
    arrange(AIC) |>
    dplyr::select(model, AIC, df.residual, fit, tidied)




```
#### Brown trout  
Most complex model has lowest AIC
```{r}
#| label: linBNT
#| 
speciesIn <- "Brown trout"
modLinearBNT <- runLinearModels(dGAM, speciesIn)

modLinearBNT |> 
    unnest(glanced) |>
    arrange(AIC) |>
    dplyr::select(model, AIC, df.residual, fit, tidied)




```
These linear models suggest that we need significant complexity in GAM models  

### GAM models for growth  
This modeling approach map() (which I would like to use), throws an error using bam(). Use gamX model approach below.
```{r}
#| label: gamModelsGrowth
#| 
runGAMModels <- function(d, speciesIn, kIn) {
  models =
    tibble(model = 
      c(
            'grWeight ~ riverGG * seasonGG * observedWeight + cohortF * ageF +             
               s(flowByRiverSR, bs = "cr", k = kIn, by = riverGG) +
               s(flowByRiverSR, bs = "cr", k = kIn, by = seasonGG) +
               s(tempSR, bs = "cr", k = kIn, by = riverGG) +
               s(tempSR, bs = "cr", k = kIn, by = seasonGG)',
            
            'grWeight ~ riverGG * seasonGG * observedWeight * tempSR * flowByRiverSR + cohortF * ageF +             
               s(flowByRiverSR, bs = "cr", k = kIn, by = riverGG) +
               s(flowByRiverSR, bs = "cr", k = kIn, by = seasonGG) +
               s(tempSR, bs = "cr", k = kIn, by = riverGG) +
               s(tempSR, bs = "cr", k = kIn, by = seasonGG)'
       )
    ) |>
    mutate(
      fit = map(model, ~ bam(.x, method = "REML", data = d |> filter(speciesGG == speciesIn, cohort %in% 2000:2013)))
  #    glanced = map(fit, broom::glance),
  #    tidied = map(fit, broom::tidy)
    )
  return(models)
}


# 
# speciesIn <- "Brook trout"
# modLinearBKT <- runGAMModels(dGAM, speciesIn, kValue)
# 
# modLinearBKT |> 
#     unnest(glanced) |>
#     arrange(AIC) |>
#     dplyr::select(model, AIC, df.residual, fit, tidied)


```

#### Brook trout - flow by river
```{r}
#| label: gamModsBKTFBR

speciesIn <- "Brook trout"
kIn <- 4

gam0 <- bam(grWeight ~ (riverGG + seasonGG + observedWeight + tempSR + flowSR)^2 + cohortF * ageF +             
               s(flowSR, bs = "cr", k = kIn, by = riverGG) +
               s(flowSR, bs = "cr", k = kIn, by = seasonGG) +
               s(tempSR, bs = "cr", k = kIn, by = riverGG) +
               s(tempSR, bs = "cr", k = kIn, by = seasonGG),
             data = dGAM |> filter(speciesGG == speciesIn, cohort %in% 2000:2013), method = "REML")
######

gam0ByRiver <- bam(grWeight ~ (riverGG + seasonGG + observedWeight + tempSR + flowByRiverSR)^2 + cohortF * ageF +             
               s(flowByRiverSR, bs = "cr", k = kIn, by = riverGG) +
               s(flowByRiverSR, bs = "cr", k = kIn, by = seasonGG) +
               s(tempSR, bs = "cr", k = kIn, by = riverGG) +
               s(tempSR, bs = "cr", k = kIn, by = seasonGG),
             data = dGAM |> filter(speciesGG == speciesIn, cohort %in% 2000:2013), method = "REML")

gam1ByRiver <- bam(grWeight ~ (riverGG + seasonGG + observedWeight + tempSR + flowByRiverSR)^2 + cohortF +             
               s(flowByRiverSR, bs = "cr", k = kIn, by = riverGG) +
               s(flowByRiverSR, bs = "cr", k = kIn, by = seasonGG) +
               s(tempSR, bs = "cr", k = kIn, by = riverGG) +
               s(tempSR, bs = "cr", k = kIn, by = seasonGG),
             data = dGAM |> filter(speciesGG == speciesIn, cohort %in% 2000:2013), method = "REML")

# age matters, what about taking out length?
gam2ByRiver <- bam(grWeight ~ (riverGG + seasonGG  + tempSR + flowByRiverSR)^2 + cohortF * ageF +             
               s(flowByRiverSR, bs = "cr", k = kIn, by = riverGG) +
               s(flowByRiverSR, bs = "cr", k = kIn, by = seasonGG) +
               s(tempSR, bs = "cr", k = kIn, by = riverGG) +
               s(tempSR, bs = "cr", k = kIn, by = seasonGG),
             data = dGAM |> filter(speciesGG == speciesIn, cohort %in% 2000:2013), method = "REML")

#nope, need to keep length
#How about without the interactions?
gam3ByRiver <- bam(grWeight ~ (riverGG + seasonGG + observedWeight + tempSR + flowByRiverSR) + cohortF * ageF +             
               s(flowByRiverSR, bs = "cr", k = kIn, by = riverGG) +
               s(flowByRiverSR, bs = "cr", k = kIn, by = seasonGG) +
               s(tempSR, bs = "cr", k = kIn, by = riverGG) +
               s(tempSR, bs = "cr", k = kIn, by = seasonGG),
             data = dGAM |> filter(speciesGG == speciesIn, cohort %in% 2000:2013), method = "REML")
#nope, need the interactions

AIC(gam0ByRiver, gam1ByRiver, gam2ByRiver, gam3ByRiver) |> arrange(AIC)
```


#### Brook trout - compare flow and flow by river gams
```{r}
#| label: gamModsCompare

predictBKTgam <- function(dIn, modInFlowByRiver){

  # predMat0 <- 
  #   expand.grid(
  #     flowSR = seq(-1.5,1.5,0.25), 
  #     tempSR = seq(-1.5,1.5,0.25), 
  #     riverGG = unique(dIn$riverGG), 
  #     seasonGG= unique(dIn$seasonGG),
  #     observedWeight = seq(2, max(dIn$observedWeight, na.rm = TRUE)/2, 20),
  #     cohortF = 2000:2013,#uunique(dIn$cohortF),
  #     ageF = 0:4,#unique(dIn$ageF),
  #     byRiver = FALSE
  # )
  # predMat <- 
  #   add_column(
  #     pred = predict.bam(modInFlow, predMat0),
  #     predMat0
  #   ) |> 
  #   rename(flow = flowSR, temp = tempSR)
  
  predMatByRiver0 <- 
    expand.grid(
      flowByRiverSR = seq(-1.5,1.5,0.25), 
      tempSR = seq(-1.5,1.5,0.25), 
      riverGG = unique(dIn$riverGG), 
      seasonGG= unique(dIn$seasonGG),
      observedWeight = seq(2, max(dIn$observedWeight, na.rm = TRUE)/2, 20),
      cohortF = 2000:2013,#unique(dIn$cohortF),
      ageF = 0:4,#unique(dIn$ageF),
      byRiver = TRUE
    ) 
  predMatByRiver <- 
    add_column(
      pred = predict.bam(modInFlowByRiver, predMatByRiver0),
      predMatByRiver0
    ) |> 
    rename(flow = flowByRiverSR, temp = tempSR)

  return(predMatByRiver)
}

gamBKTPredictions <- predictBKTgam(dGAM, gam0ByRiver)

```

#### Graph predictions
```{r}
#| label: graphGAMPredictions
unique(gamBKTPredictions$observedWeight)

ggplot(
  gamBKTPredictions |> 
    filter(
      observedWeight %in% c(22), 
      cohortF == 2012, 
      ageF == 0
    ), 
  aes(temp, pred, color = (flow), group = (flow))) +
  geom_point(alpha = 0.2) +
  labs(x = "Stream temperature", y = "Predicted growth rate") +
  #geom_smooth(se = FALSE) +
  geom_hline(yintercept = 0) +
  facet_grid(seasonGG~riverGG+byRiver, scales = 'free')

```


## Logistic models for environmental effects on propNeg

### Standardize flow and temperature

```{r}
#| label: flowTempZ
#| 
# getZ <- function(d){
#   list(
#     mean = mean(d, na.rm = TRUE),
#     sd = sd(d, na.rm = TRUE)
#   )
# }
# fZ <- getZ(cd1$meanFlowByRiver)
# fT <- getZ(cd1$meanTemperature)

z <- cd1 |> 
  group_by(season) |> 
  summarize(fZMean = mean(meanFlowByRiver, na.rm = TRUE),
            fZSd = sd(meanFlowByRiver, na.rm = TRUE),
            fTMean = mean(meanTemperature, na.rm = TRUE),
            fTSd = sd(meanTemperature, na.rm = TRUE)
            )

cd1 <- cd1 |> 
  left_join(z) |> 
  mutate(meanFlowByRiverZ = (meanFlowByRiver - fZMean)/fZSd,
         meanTempZ = (meanTemperature - fTMean)/fTSd)
```

### Raw data with logistic fit for flow effect

```{r}
#| label: logisticFlow

ggplot(cd1 |> filter(!is.na(negGrowth)), aes(meanFlowByRiverZ, negGrowth * 1, color = river)) +
  geom_point(alpha = 0.1) +
  geom_smooth(method = "glm", 
              method.args = list(family = "binomial"),
              se = FALSE) +
  labs(y = "Probability of negative growth in mass", x = "Mean flow") +
  facet_grid(season~species)

ggplot(cd1 |> filter(!is.na(negGrowth)), aes(meanFlowByRiverZ, negGrowth * 1, color = river)) +
  geom_point(alpha = 0.1) +
  geom_smooth(method = "glm", 
              method.args = list(family = "binomial"),
              se = FALSE) +
  labs(y = "Probability of negative growth in mass", x = "Mean flow") +
  facet_grid(season~species+river)
```

### Raw data with logistic fit for temperature effect

```{r}
#| label: logisticTemp

ggplot(cd1 |> filter(!is.na(negGrowth), meanTempZ > -4), aes(meanTempZ, negGrowth * 1, color = river)) +
  geom_point(alpha = 0.1) +
  geom_smooth(method = "glm", 
              method.args = list(family = "binomial"),
              se = FALSE) +
  labs(y = "Probability of negative growth in mass", x = "Mean temperature") +
  facet_grid(season~species)
```

## Logistic model for flow and temperature, one species at a time

Species occupy different rivers, so an all-species model will have
unequal river representation across species and don't need to predict
where species are not.

Need to think through whether these models make sense...

```{r}
#| label: runLogisticModels
runSppModels <- function(d, speciesIn) {
  models =
    tibble(model = 
      c(
          'negGrowth ~ meanFlowByRiverZ * meanTempZ * river',
          'negGrowth ~ meanFlowByRiverZ * meanTempZ + river',
          'negGrowth ~ meanFlowByRiverZ + meanTempZ + river', 
          
          'negGrowth ~ meanFlowByRiverZ * meanTempZ' , 
          
          'negGrowth ~ meanFlowByRiverZ * meanTempZ * river * factor(season)',
          
          'negGrowth ~ meanFlowByRiverZ + meanTempZ * river * factor(season)', 
          
          'negGrowth ~ meanFlowByRiverZ * meanTempZ * river * factor(season) * observedWeight' 
        )
    ) |>
    mutate(
      fit = map(model, ~glm(.x, family = "binomial", data = d |> filter(species == speciesIn))),
      glanced = map(fit, broom::glance),
      tidied = map(fit, broom::tidy)
    )
  return(models)
}

```

### Predict p(neg growth in mass)

```{r}
#| label: predictLogistic

predictSpp <- function(dIn, modIn, speciesIn){
  d <- dIn |> filter(species == speciesIn)
  
  byAIC = modIn |> 
    unnest(glanced) |>
    arrange(AIC) |>
    dplyr::select(model, AIC, df.residual, fit, tidied)
  
  predDSpp <- expand.grid(
    meanFlowByRiverZ = seq(-1.5,1.5,0.25), 
    meanTempZ = seq(-1.5,1.5,0.25), 
    river = unique(d$river), 
    season = unique(d$season),
    observedWeight = seq(min(d$observedWeight, na.rm = TRUE), max(d$observedWeight, na.rm = TRUE), 40)
    )
  
  
  pred <- add_column(predDSpp, 
                        pred = predict.glm(byAIC[1,]$fit[[1]], predDSpp, type = "response")) |> 
             mutate(groupT = paste(river, season, meanTempZ, sep = "_"),
                    groupF = paste(river, season, meanFlowByRiverZ, sep = "_"))
  return(
    list(
      pred = pred,
      byAIC = byAIC
      )
  )
}   


```

### Brook trout

```{r}
#| label: modBKT
speciesIn <- "bkt"
modBKT <- runSppModels(cd1, speciesIn)
predBKT <- predictSpp(cd1, modBKT, speciesIn)


ggplot(predBKT$pred |> filter(meanTempZ %in% c(-1.5,0,1.5), meanFlowByRiverZ %in% c(-1.5,0,1.5)), aes(observedWeight, pred, group = groupF)) +
  geom_line(aes(color = factor(meanFlowByRiverZ))) +
  labs(y = "Probability of negative growth in mass", x = "Body mass (mg)") +
  facet_grid(season~river+meanTempZ)

```

### Brown trout

```{r}
#| label: modBnt
#| 
speciesIn <- "bnt"

modBNT <- runSppModels(cd1, speciesIn)
predBNT <- predictSpp(cd1, modBNT, speciesIn)


ggplot(predBNT$pred |> filter(meanTempZ %in% c(-1.5,0,1.5), meanFlowByRiverZ %in% c(-1.5,0,1.5)), aes(observedWeight, pred, group = groupF)) +
  geom_line(aes(color = factor(meanFlowByRiverZ))) +
  labs(y = "Probability of negative growth in mass", x = "Body mass (mg)") +
  facet_grid(season~river+meanTempZ)
```

### Atlantic salmon

```{r}
#| label: modAts
#| 
speciesIn <- "ats"

modATS <- runSppModels(cd1, speciesIn)
predATS <- predictSpp(cd1, modATS, speciesIn)


ggplot(predATS$pred |> filter(meanTempZ %in% c(-1.5,0,1.5), meanFlowByRiverZ %in% c(-1.5,0,1.5)), aes(observedWeight, pred, group = groupF)) +
  geom_line(aes(color = factor(meanFlowByRiverZ))) +
  labs(y = "Probability of negative growth in mass", x = "Body mass (mg)") +
  facet_grid(season~river+meanTempZ)
```

## Two growth models

### Size-adjusted growth model

#### Get l-w relationships for size-adjusted growth model

Here, we will explore size-independent growth in mass, see
`Sigourney, D. B., B H Letcher, M. Obedzinski, and R. A. Cunjak. Size-Independent Growth in Fishes: Patterns, Models and Metrics. Journal of Fish Biology 72, no. 10 (2008): 243555. https://doi.org/10.1111/j.1095-8649.2008.01830.x.`
Size-independent growth (`grWeightS`) is calculated in
getDataElectro_targets.R file using addSizeIndGrowthWeight().\
`grWeightS` uses the slope of `log(observedWeight)` \~ `log(grWeight)`
for each river and season and species to adjust growth.\
A simple linear models shows we need to get separate slopes for each
river and season and species.

```{r}
#| label: grW_models
ggplot(cd1 |> filter(grWeight > 0), 
       aes(log(observedWeight), log(grWeight), color = riverGG)) +
  geom_point(alpha = 0.05)  +
  geom_smooth(method = "lm", se = FALSE) +
  labs(y = "Log of mass (mg)", x = "Log of growth in mass") +
  scale_color_discrete() +
  facet_grid(seasonGG~speciesGG)

mod0 <- lm(log(grWeight) ~ log(observedWeight), data = cd1 |> filter(grWeight > 0))
mod1 <- lm(log(grWeight) ~ log(observedWeight) * river, data = cd1 |> filter(grWeight > 0))
mod2 <- lm(log(grWeight) ~ log(observedWeight) * river * factor(season), data = cd1 |> filter(grWeight > 0))
mod3 <- lm(log(grWeight) ~ log(observedWeight) * factor(season), data = cd1 |> filter(grWeight > 0))
mod4 <- lm(log(grWeight) ~ log(observedWeight) * river * factor(season) * species, data = cd1 |> filter(grWeight > 0))

AIC(mod0,mod1,mod2,mod3,mod4) |> arrange(AIC)
```

### Compare size-adjusted and instantaneous growth

Differences don't seem big enough to warrant the extra complexity from
the s-adjusted growth metric

```{r}
#| label: grWeight_v_grWeightS
ggplot(cd1, aes(grWeight, grWeightS, color = riverGG)) +
  geom_point(alpha = 0.05)  +
  geom_smooth(method = "lm", se = FALSE) +
  geom_abline(intercept = 0, slope = 1) +
  labs(y = "Gowth in mass", x = "Size-adjusted growth in mass") +
  scale_color_discrete() +
  facet_grid(seasonGG~speciesGG)

ggplot(cd1 |> 
         dplyr::select(observedWeight, grWeight, grWeightS, seasonGG, speciesGG) |> 
         pivot_longer(cols = c(grWeight, grWeightS)),
       
       aes(log(observedWeight), value, color = name)) +
  geom_point(alpha = 0.05 )  +
  geom_smooth(method = "lm", se = FALSE) +
  labs(y = "Log of mass (mg)", x = "Growth in mass") +
  scale_color_discrete() +
  facet_grid(seasonGG~speciesGG)
```

## Compare growth in length and mass

### Growth rate in length vs size-independent growth rate in mass

```{r}
#| label: grLen_v_grWeightS
ggplot(cd1, aes(grLength, grWeightS)) +
  geom_point(aes(size = observedWeight), alpha = 0.02) +
  guides(size = guide_legend(override.aes = list(alpha = 1))) +
  labs(x = "Growth in length", y = "Size-adjusted growth in mass") +
  scale_color_discrete() +
  facet_grid(seasonGG~speciesGG)
```

## **Deprecated** logistic models below

### Logistic model for flow and temperature, all species

### Model selection

```{r}
#| label: logistic models all species
modSpecies <- c("bkt", "ats", "bnt")
cd1Mod <- cd1 |> filter(species == modSpecies)

modFT0 <- glm(negGrowth ~ meanFlowByRiverZ * meanTempZ * river * species, family = "binomial", data = cd1Mod)
modFT1 <- glm(negGrowth ~ meanFlowByRiverZ * meanTempZ * river + species, family = "binomial", data = cd1Mod)
modFT2 <- glm(negGrowth ~ meanFlowByRiverZ * meanTempZ + river + species, family = "binomial", data = cd1Mod)
modFT3 <- glm(negGrowth ~ meanFlowByRiverZ + meanTempZ + river + species, family = "binomial", data = cd1Mod)
modFT4 <- glm(negGrowth ~ meanFlowByRiverZ * meanTempZ + river , family = "binomial", data = cd1Mod)
modFT5 <- glm(negGrowth ~ meanFlowByRiverZ * meanTempZ + river , family = "binomial", data = cd1Mod)
modFT6 <- glm(negGrowth ~ meanFlowByRiverZ * meanTempZ + species , family = "binomial", data = cd1Mod)
modFT7 <- glm(negGrowth ~ meanFlowByRiverZ * meanTempZ, family = "binomial", data = cd1Mod)


modFT8 <- glm(negGrowth ~ meanFlowByRiverZ * meanTempZ * river * factor(season) + species, family = "binomial", data = cd1Mod)
modFT9 <- glm(negGrowth ~ meanFlowByRiverZ * meanTempZ * river * factor(season) * species, family = "binomial", data = cd1Mod)
modFT10 <- glm(negGrowth ~ meanFlowByRiverZ + meanTempZ * river * factor(season) * species, family = "binomial", data = cd1Mod)

modFT11 <- glm(negGrowth ~ meanFlowByRiverZ * meanTempZ * river * factor(season) * observedWeight + species, family = "binomial", data = cd1Mod)
modFT12 <- glm(negGrowth ~ meanFlowByRiverZ * meanTempZ * river * factor(season) * species * observedWeight, family = "binomial", data = cd1Mod)

AIC(modFT0, modFT1,modFT2, modFT3,modFT4,modFT5, modFT6,modFT7,modFT8,modFT9,modFT10, modFT11, modFT12) |> arrange(AIC)
```

### Model results for additive model (not very good)

```{r}
#| label: badAdditiveModel
# https://www.statology.org/logistic-regression-in-r/
#summary(modFT3)
#anova(modFT3)

pscl::pR2(modFT3)["McFadden"]
caret::varImp(modFT3) |> arrange(desc(Overall))


```

### Model results for full interactive model (best model)

```{r}
#| label: fullModel
# https://www.statology.org/logistic-regression-in-r/
#summary(modFT11)
#anova(modFT11)

# full interaction
pscl::pR2(modFT11)["McFadden"]
caret::varImp(modFT11) |> arrange(desc(Overall))
```

### Precict p(neg growth in mass)

```{r}
#| label: predictFull
#| 
predD <- expand.grid(
  meanFlowByRiverZ = seq(-1.5,1.5,0.25), 
  meanTempZ = seq(-1.5,1.5,0.25), 
  river = unique(cd1$river), 
  species = unique(cd1$species), 
  season = unique(cd1$season),
  observedWeight = seq(min(cd1$observedWeight, na.rm = TRUE), max(cd1$observedWeight, na.rm = TRUE), 40)
  )


pred <- add_column(predD, 
                   pred = predict.glm(modFT11, predD, type = "response")) |> 
        mutate(groupT = paste(river, species, season, meanTempZ, sep = "_"),
               groupF = paste(river, species, season, meanFlowByRiverZ, sep = "_"))
                   
```

### Plot predictions

```{r}
#| label: plotPred

ggplot(pred |> filter(meanTempZ %in% c(-1.5,0,1.5), observedWeight == 161.3), aes(meanFlowByRiverZ, pred, group= groupT)) +
  geom_line(aes(color = factor(meanTempZ))) +
  labs(y = "Probability of negative growth in mass", x = "Mean flow") +
  facet_grid(season~species+river)
```
