# Condition factor

Explore relative condition factor which is the ratio between mass and length, where length is raised to the power of the exponent of the log10(mass):log10(length) relationship. Need to check whether the mass:length relationship varies among species and, if so, use species-specific exponents.

```{r}
#| label: dataModelCFptions
#| include: false
knitr::opts_chunk$set(message = FALSE, warning = FALSE)
```

```{r}
#| label: librariesModelsCF
#| echo: false
library(getWBData)
library(lubridate)
library(kableExtra)
library(GGally)
library(lme4)
library(targets)
library(relaimpo)
library(caret)
library(pscl)
library(tidyverse)

`%notin%` <- Negate(`%in%`)
```

## Get data

### Get electrofishing data

```{r}
#| label: readInAllDataCF
#| cache: false
library(targets)

cd <- tar_read(cdWB_electro_target) # raw data
cd1 <- tar_read(cd1_target)

relCF_byYear <- tar_read(relCF_byYear_target)
relCF_byYearWide <- tar_read(relCF_byYearWide_target)
```

### Plot raw lengths/mass

```{r}
#| label: lwAll

ggplot(cd1, aes(observedLength, observedWeight, color = speciesGG)) +
  geom_point(alpha = 0.1) +
  guides(color = guide_legend(override.aes = list(alpha = 1))) +
  scale_x_log10("Length (mm)") +
  scale_y_log10("Mass (g)") +
  scale_color_discrete() +
  facet_wrap(~speciesGG) 
  

ggplot(cd1, aes(observedLength, observedWeight, color = speciesGG)) +
  geom_point(alpha = 0.05) +
  geom_smooth(method = "lm") +
  scale_x_log10("Length (mm)") +
  scale_y_log10("Mass (g)") +
  scale_color_discrete()
```

### Length/mass regression

Does species matter? **Yes** Differences appear very small on the graphs, but there are significant differences in slope between brook trout and the other two species (brown trout and Atlantic salmon).\
Use slopes for each species in relative condition factor calculation. Model code below is repeated in `getDataElectro_targets.R` in `addCF()` to add `cf` and `relCF` to `cdWB_electro_target`

```{r}
mod1 <- lm(log10(observedWeight) ~ log10(observedLength) * factor(species), data = cd1)
mod2 <- lm(log10(observedWeight) ~ log10(observedLength) + factor(species), data = cd1)
mod3 <- lm(log10(observedWeight) ~ log10(observedLength), data = cd1)

AIC(mod1,mod2,mod3) |> arrange(AIC)
summary(mod1)

(CF_coefs <- tibble(
  species = c("ats", "bkt", "bnt"),
  intercept = c(
    coef(mod1)[["(Intercept)"]],
    coef(mod1)[["(Intercept)"]] + coef(mod1)[["factor(species)bkt"]],
    coef(mod1)[["(Intercept)"]] + coef(mod1)[["factor(species)bnt"]]
  ),
  slope = c(
    coef(mod1)[["log10(observedLength)"]],
    coef(mod1)[["log10(observedLength)"]] + coef(mod1)[["log10(observedLength):factor(species)bkt"]],
    coef(mod1)[["log10(observedLength)"]] + coef(mod1)[["log10(observedLength):factor(species)bnt"]]
  )
))
```

### Relative condition factor by species

```{r}
#| label: relCF
ggplot(cd1, aes(relCF, color = speciesGG)) +
  geom_freqpoly(bins = 50) +
  labs(x = "Relative condition factor", y = "Frequency") +
  scale_color_discrete() +
  facet_wrap(~seasonGG)
```

### Relative condition factor by species across years

```{r}
#| label: relCF_years
ggplot(relCF_byYear, aes(year, mRCF, color = speciesGG)) +
  geom_point() +
  geom_line() +
  scale_color_discrete() +
  labs(x = "Year", y = "Relative condition factor") +
  facet_grid(riverGG~seasonGG)
```

### Relative condition factor species comparison
Do relative condition factors covary between species?
```{r}
#| label: relCF_yearsWide

ggplot(relCF_byYearWide |> filter(riverGG != "Isolated Small"), aes(`Brook trout`, `Brown trout`)) +
  geom_point(aes(color = factor(year))) +
  geom_smooth(method = "lm", se = FALSE, color = 'darkgrey') +
  scale_color_discrete() +
  labs(x = "Brook trout relative condition factor", y = "Brown trout relative condition factor") +
  facet_grid(riverGG~seasonGG)

ggplot(relCF_byYearWide |> filter(riverGG == "West Brook"), aes(`Brook trout`, `Atlantic salmon`)) +
  geom_point(aes(color = factor(year))) +
  geom_smooth(method = "lm", se = FALSE, color = 'darkgrey') +
  scale_color_discrete() +
  labs(x = "Brook trout relative condition factor", y = "Atlantic salmon relative condition factor") +
  facet_wrap(~seasonGG)

ggplot(relCF_byYearWide |> filter(riverGG == "West Brook"), aes(`Brown trout`, `Atlantic salmon`)) +
  geom_point(aes(color = factor(year))) +
  geom_smooth(method = "lm", se = FALSE, color = 'darkgrey') +
  scale_color_discrete() +
  labs(x = "Brown trout relative condition factor", y = "Atlantic salmon relative condition factor") +
  facet_wrap(~seasonGG)
```

### Seasonal comparison
TO ADD